### 2017-11-11 voltagebudget/analysis/sfn_figures.Rmd
SFN poster final

---
title: "Exp1 - Opt w_max and {A,phi)"
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}
```

# Load and process

```{r, message=FALSE}
results <- read_csv("../data/sfn10.csv")
```

```{r, fig.width=1.5, fig.height=1}
results %>% 
  # filter(computation_scores < 0.6, communication_scores > 0) %>%
  ggplot(aes(y=computation_voltages*1e4, x=communication_voltages*1e2, color=As)) +
  scale_color_gradient(low="mediumpurple", high="red3", name="A") +
  xlab("Vo (mvolt)") + ylab("Vc (mvolt)") +
  geom_point(alpha=0.3) + 
  theme_classic()
```

```{r, fig.width=.85, fig.height=.75}
results %>% 
  # filter(computation_scores < 0.6, communication_scores > 0) %>%
  ggplot(aes(x=1e2*communication_voltages, y=communication_scores)) +
  geom_point(color="red3", alpha=0.6) + 
  xlab("Vo (mvolt)") +
  ylab("C (counts)") +
  ylim(0, 20) +
  theme_classic() -> p1
print(p1)
```

```{r, fig.width=.85, fig.height=.75}
results %>% 
  ggplot(aes(x=1e4*computation_voltages, y=30e-3-computation_scores)) +
  geom_point(color="mediumpurple", alpha=0.6) + 
  xlab("Vc (mvolt)") +
  ylab("S (sec)") +
  ylim(0, 0.030) +
  theme_classic() -> p2
print(p2)
```

```{r, fig.width=.95, fig.height=.75}
results %>% 
  filter(computation_scores < 0.6, communication_scores > 0) %>%
  ggplot(aes(x=20+1e2*communication_voltages + 1e4*computation_voltages, y=normalize(communication_scores))) +
  geom_smooth(method = "glm",  method.args = list(family = "binomial"), color="red3", alpha=0.5, se=FALSE) + 
  geom_point(color="red3", alpha=0.2) + 
  xlab("Vf (mvolt)") +
  ylab("Norm. scores") +
  geom_smooth(aes(x=20+1e2*communication_voltages + 1e4*computation_voltages, y=normalize(30e-3-computation_scores)), 
              method = "glm",  method.args = list(family = "binomial"), color="mediumpurple", se=FALSE) + 
  geom_point(aes(x=20+1e2*communication_voltages+ 1e4*computation_voltages, y=normalize(30e-3-computation_scores)), color="mediumpurple", alpha=0.2) + 
  # geom_vline(xintercept = 8.8373, color="black")+
  theme_classic() -> p3
print(p3)
```

```{r, fig.width=.95, fig.height=2}
grid.arrange(p1, p2, p3, ncol=1)

```