### 2018-01-28 voltagebudget/analysis/ranked_sum_analysis.Rmd
cleaned up Rmd plotting

---
title: "General analysis notebook for `forward` budget models."
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}
```

# Load and process a series of ranks.

```{r, message=FALSE, fig.width=3, fig.height=1}
# Load min, max V_free and neighbors
exp_name <- "exp18"
freq = "30"

ranked <- NULL

rank <- "1"
name <- paste("../data/", exp_name, "_d-2e-3_f", freq, "_r", rank, ".csv", sep="")
rdata <- read_csv(paste(name, sep=""))
rdata$rank <- rep(rank, nrow(rdata))
ranked <- rbind(ranked, rdata)

# rank <- "2"
# name <- paste("../data/", exp_name, "_d-2e-3_f", freq, "_r", rank, ".csv", sep="")
# rdata <- read_csv(paste(name, sep=""))
# rdata$rank <- rep(rank, nrow(rdata))
# ranked <- rbind(ranked, rdata)

rank <- "3"
name <- paste("../data/", exp_name, "_d-2e-3_f", freq, "_r", rank, ".csv", sep="")
rdata <- read_csv(paste(name, sep=""))
rdata$rank <- rep(rank, nrow(rdata))
ranked <- rbind(ranked, rdata)

rank <- "50"
name <- paste("../data/", exp_name, "_d-2e-3_f", freq, "_r", rank, ".csv", sep="")
rdata <- read_csv(paste(name, sep=""))
rdata$rank <- rep(rank, nrow(rdata))
ranked <- rbind(ranked, rdata)

# rank <- "49"
# name <- paste("../data/", exp_name, "_d-2e-3_f", freq, "_r", rank, ".csv", sep="")
# rdata <- read_csv(paste(name, sep=""))
# rdata$rank <- rep(rank, nrow(rdata))
# ranked <- rbind(ranked, rdata)

rank <- "48"
name <- paste("../data/", exp_name, "_d-2e-3_f", freq, "_r", rank, ".csv", sep="")
rdata <- read_csv(paste(name, sep=""))
rdata$rank <- rep(rank, nrow(rdata))
ranked <- rbind(ranked, rdata)

ranked$rank <- as.numeric(ranked$rank)

# -----------------------------------------------------
# TODO: manual color assignment, setting alphas to rank?
ranked %>%
  filter(N!=2) %>% 
  ggplot(aes(x=N, y=sort(errors), color=rank, group=rank)) +
  geom_line(alpha=0.4) +
  scale_color_gradient(low="darkolivegreen3", high="goldenrod2") +
  theme_classic() -> p3

ranked %>%
  filter(N!=2) %>% 
  ggplot(aes(x=N, y=sort(variances), color=rank, group=rank)) +
  geom_line(alpha=0.4) +
  scale_color_gradient(low="darkolivegreen3", high="goldenrod2") +
  theme_classic() -> p4

grid.arrange(p3, p4, nrow=1)
```