### 2018-01-30 voltagebudget/analysis/ranked_analysis.Rmd

---
title: "General analysis notebook for `forward` budget models."
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}
```

# Load and process indivdual ranks

```{r, message=FALSE}
name <- "../data/exp18_d-2e-3_f12_r50.csv"
results <- read_csv(paste(name, sep=""))
```

```{r, fig.width=2.5, fig.height=1.5}
results %>% 
  ggplot(aes(x=delta_variances)) +
  geom_histogram(bins=20) + 
  xlab("Variance") +
  theme_classic() -> p1

results %>% 
  ggplot(aes(x=errors)) +
  geom_histogram(bins=20) + 
  xlab("Error") +
  theme_classic() -> p2

results %>% 
  ggplot(aes(x=V_osc)) +
  geom_histogram(bins=20) +
  xlab("Vc (volt)") +
  theme_classic() -> p3

results %>% 
  ggplot(aes(x=V_comp)) +
  geom_histogram(bins=20) +
  xlab("Vo (volt)") +
  theme_classic() -> p4

grid.arrange(p1, p2, p3, p4, nrow=2)
```

```{r, fig.width=2.5, fig.height=1}
results %>% 
  ggplot(aes(x=V_free, y=V_osc, color=errors)) +
  geom_point() + 
  theme_classic()
```

```{r, fig.width=2.5, fig.height=1}
results %>% 
  ggplot(aes(x=N, y=sort(errors))) +
  ylim(0, 0.030) +
  geom_point() + 
  theme_classic()
```

```{r, fig.width=2.5, fig.height=1}
results %>% 
  ggplot(aes(x=N, y=sort(delta_variances))) +
  geom_point() + 
  ylim(-0.030, 0.030) +
  geom_hline(aes(yintercept=0), color="black", alpha=0.6) +
  theme_classic()
```

```{r, fig.width=1.5, fig.height=1}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=V_comp, y=normalize(delta_variances))) +
  geom_point(color="red", alpha=0.5) + 
  xlab("Vc (mvolt)") +
  ylab("Rel. scores (AU)") +
  geom_point(aes(x=V_comp, y=normalize(errors)), color="purple") + 
  theme_classic()
```