### 2018-02-23 voltagebudget/analysis/ranked_analysis.Rmd
vis updates

---
title: "General analysis notebook for `forward` budget models."
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}
```

# Load and process indivdual ranks

```{r, message=FALSE}
name <- "../data/exp44_d-0.2e-3_f12_r300.csv"
results <- read_csv(paste(name, sep=""))
print(range(results$rank))
results %>%
  filter(n_spikes > 0, rank==299) -> results
```

```{r, fig.width=2.5, fig.height=1.5}
results %>% 
  ggplot(aes(x=variances)) +
  geom_histogram(bins=20) + 
  xlab("Variance") +
  theme_classic() -> p1

results %>% 
  ggplot(aes(x=errors)) +
  geom_histogram(bins=20) + 
  xlab("Error") +
  theme_classic() -> p2

results %>% 
  ggplot(aes(x=V_osc)) +
  geom_histogram(bins=20) +
  xlab("Vc (volt)") +
  theme_classic() -> p3

results %>% 
  ggplot(aes(x=V_comp)) +
  geom_histogram(bins=20) +
  xlab("Vo (volt)") +
  theme_classic() -> p4

grid.arrange(p1, p2, p3, p4, nrow=2)
```

## A and V

```{r, fig.width=2.5, fig.height=1}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=sort(V_free), y=errors, color=As)) +
  geom_point() + 
  theme_classic()
```

```{r, fig.width=2.5, fig.height=1}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=As, y=V_osc/V_b)) +
  geom_point() + 
  theme_classic()
```

```{r, fig.width=2.5, fig.height=1}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=V_osc, y=V_comp, color=As)) +
  geom_point() + 
  theme_classic()
```

### As

```{r, fig.width=2.5, fig.height=1}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=As, y=variances)) +
  geom_point() + 
  theme_classic()
```

```{r, fig.width=2.5, fig.height=1}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=As, y=errors)) +
  geom_point() + 
  theme_classic()
```

```{r, fig.width=2.5, fig.height=1}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=As, y=n_spikes)) +
  geom_point() + 
  theme_classic()
```

### V_f

```{r, fig.width=2.5, fig.height=1}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=sort(V_free), y=variances)) +
  geom_point() + 
  theme_classic()
```

### V_osc/V_B

```{r, fig.width=2.5, fig.height=1}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=V_osc, y=variances)) +
  geom_point() + 
  theme_classic()
```

```{r, fig.width=2.5, fig.height=1.5}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=V_osc/V_b, y=errors)) +
  geom_point() + 
  theme_classic()
```

```{r, fig.width=2.5, fig.height=1.5}
results %>% 
  # filter(errors < 0.6, coincidence_count > 0) %>%
  ggplot(aes(x=errors, y=variances, color=As)) +
  geom_point() + 
  theme_classic()
```