### 2018-03-02 voltagebudget/analysis/min_max_delay_analysis.Rmd
tweaks to vis

---
title: "min/max V_f - delay (d) experiment analysis."
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}
```

# Load and process

```{r, message=FALSE}
exp <- "62"
freq <- c("20")

delays <- c("-2e-3", "-4e-3", "-6e-3", "-8e-3", "-10e-3")

results <- NULL
for(d in delays){
  name <- paste("../data/exp", exp, "_d", d, "_f", freq, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  df %>%
    filter(n_spikes > 0) -> df

  df$delay <- rep(d, nrow(df))

  results <- rbind(results, df)
}

results$delay <- factor(results$delay, levels=delays)
rm(df, exp, delays, d, name, freq)
```

# Overall distributions

```{r, fig.width=2.5, fig.height=1.5}
results %>% 
  ggplot(aes(x=variances)) +
  geom_histogram(bins=20) + 
  xlab("Variance") +
  theme_classic() -> p1

results %>% 
  ggplot(aes(x=errors)) +
  geom_histogram(bins=20) + 
  xlab("Error") +
  theme_classic() -> p2

results %>% 
  ggplot(aes(x=V_osc/V_b)) +
  geom_histogram(bins=20) +
  xlab("Vc (volt)") +
  theme_classic() -> p3

results %>% 
  ggplot(aes(x=V_comp/V_b)) +
  geom_histogram(bins=20) +
  xlab("Vo (volt)") +
  theme_classic() -> p4

grid.arrange(p1, p2, p3, p4, nrow=2)
rm(p1, p2, p3, p4)
```

# As

```{r, fig.width=1.2, fig.height=2.5}
results %>% 
  ggplot(aes(x=As*1e9, y=n_spikes, color=delay)) +
  geom_line() + 
  theme_classic() +
  facet_grid(delay~.) +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=As*1e9, y=variances, color=delay)) +
  facet_grid(delay~.) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

- why is var so variable?

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=As*1e9, y=errors, color=delay)) +
  facet_grid(delay~.) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=V_comp/V_b, y=errors, color=delay)) +
  facet_grid(delay~.) +
  geom_point(size=0.1) + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.25, fig.height=3}
results %>% 
  ggplot(aes(x=variances, y=errors, color=delay)) +
  geom_line() + 
  theme_classic() +
  facet_grid(delay~.) +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=n_spikes, y=variances, color=delay)) +
  geom_point(size=0.1) + 
  theme_classic() +
  facet_grid(delay~.) +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=n_spikes, y=errors, color=delay)) +
  geom_point(size=0.1) + 
  theme_classic() +
  facet_grid(delay~.) +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

# Budget

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=As*1e9, y=V_osc/V_b, color=delay)) +
  geom_line() + 
  theme_classic() +
  facet_grid(delay~.) +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=As*1e9, y=V_comp/V_b, color=delay)) +
  geom_line() + 
  theme_classic() +
  facet_grid(delay~.) +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=V_osc/V_b, y=V_comp/V_b, color=delay)) +
  geom_line() +
  theme_classic() +
  facet_grid(delay~.) +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=V_osc/V_b, y=n_spikes, color=delay)) +
  geom_line() + 
  theme_classic() +
  facet_grid(delay~.) +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=V_osc/V_b, y=variances, color=delay)) +
  geom_line() + 
  theme_classic() +
  facet_grid(delay~.) +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.25, fig.height=2.5}
results %>% 
  ggplot(aes(x=V_osc/V_b, y=errors, color=delay)) +
  geom_line() + 
  theme_classic() +
  facet_grid(delay~.) +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```