### 2018-03-10 voltagebudget/analysis/sweep_A_analysis.Rmd
Update sweep_A vis

---
title: "Sweep A - general experiment analysis."
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}
```

# Load and process

```{r, message=FALSE}
exp <- "71"
freqs <- c("8", "12", "20", "24", "30")

results <- NULL
for(freq in freqs){
  name <- paste("../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  df %>%
    filter(n_spikes > 0) -> df

  df$freq <- rep(freq, nrow(df))
  results <- rbind(results, df)
}

results$freq <- factor(results$freq, levels=freqs)
rm(df, exp, freq, freqs, name)
```

# Overall distributions

```{r, fig.width=2.5, fig.height=1.5}
results %>% 
  ggplot(aes(x=variances)) +
  geom_histogram(bins=20) + 
  xlab("Variance") +
  theme_classic() -> p1

results %>% 
  ggplot(aes(x=errors)) +
  geom_histogram(bins=20) + 
  xlab("Error") +
  theme_classic() -> p2

results %>% 
  ggplot(aes(x=V_osc/V_b)) +
  geom_histogram(bins=20) +
  xlab("Vc (volt)") +
  theme_classic() -> p3

results %>% 
  ggplot(aes(x=V_comp/V_b)) +
  geom_histogram(bins=20) +
  xlab("Vo (volt)") +
  theme_classic() -> p4

grid.arrange(p1, p2, p3, p4, nrow=2)
rm(p1, p2, p3, p4)
```

# As

```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  filter(rank == 0) %>% 
  ggplot(aes(x=As, y=n_spikes_pop, color=freq)) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  filter(rank == 0) %>% 
  ggplot(aes(x=As, y=variances_pop, color=freq)) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.6, fig.height=1.6}
results %>% 
  filter(rank == 0) %>% 
  ggplot(aes(x=As, y=errors_pop, color=freq)) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  filter(rank == 0) %>% 
  ggplot(aes(x=variances_pop, y=errors_pop, color=freq)) +
  geom_point(size=.05) + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.5, fig.height=3.5}
results %>% 
  ggplot(aes(x=variances_pop, y=errors_pop, color=freq)) +
  geom_point(size=.05) + 
  facet_grid(freq~.) +
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

# Number spikes

```{r, fig.width=1.6, fig.height=1.6}
results %>% 
  filter(rank == 0) %>% 
  ggplot(aes(x=n_spikes_pop, y=variances_pop, color=freq)) +
  geom_point(size=0.1) + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.6, fig.height=1.6}
results %>% 
  filter(rank == 0) %>% 
  ggplot(aes(x=n_spikes_pop, y=errors_pop, color=freq)) +
  geom_point(size=0.1) + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

# Budget

```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  ggplot(aes(x=As*1e9, y=V_osc, color=freq)) +
  geom_point() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

- TODO: port over Vo/Vc plots....