### 2018-03-23 voltagebudget/analysis/optimal_analysis.Rmd
Added optimal analysis Rmb

---
title: "Optimal coordination - general experiment analysis."
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(readr)
library(scales)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}
```

# Load and process

```{r, message=FALSE}
# Data
algs <- c("max", "left", "uni", "cc")
alg_names <- c("max", "left", "uniform", "coincidence")
exps <- c(82, 83, 84, 85)

results <- NULL
for (n in 1:4){
  df <- read_csv(paste("../data/exp", as.character(exps[n]), "_", algs[n], ".csv", sep=""))
  df$alg <- rep(alg_names[n], nrow(df))

  results <- rbind(results, df)
}

# Spikes
# Load ref spikes
ref <- read_csv("../data/stim1.csv")
ref$set <- rep("reference", nrow(ref))

# Load exp spikes
exps <- c(86, 87, 88, 89)
spikes <- NULL
for (n in 1:4){
  df <- read_csv(paste("../data/exp", as.character(exps[n]), "_", algs[n], "_c0.2.csv", sep=""))
  df$set <- rep("experiment", nrow(df))

  # Add ref
  df <- rbind(df, ref)

  df$alg <- rep(alg_names[n], nrow(df))
  spikes <- rbind(spikes, df)
}

rm(df, n, ref)
```

```{r, fig.width=1.4, fig.height=0.8}
results %>% 
  filter(alg != "coincidence") %>%
  ggplot(aes(x=errors*1e3, y=obs_variances*1e3, linetype=alg)) +
  geom_line(size=0.6, alpha=0.9) +
  scale_linetype_manual(values=c("longdash", "solid", "dotdash"), name="Algorithm") +
  theme_classic() +
  labs(x="Error (ms)", y="Variance (ms)")
```

```{r, fig.width=1.0, fig.height=1.5}
spikes %>% 
  filter(alg != "coincidence") %>%
  ggplot(aes(x=ts, y=ns, color=set)) +
  geom_point(size=0.1) +
  facet_grid(alg~.) +
  scale_color_manual(values=c("black", "grey")) +
  theme_classic() +
  theme(
    legend.position="none", 
    legend.key.size = unit(.4, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8))
```