### 2018-04-04 voltagebudget/analysis/weight_analysis.Rmd
Added weight analysis Rmd

---
title: "Sweep Z - general experiment analysis."
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}
data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-mad(x)
   ymax <- m+mad(x)
   return(c(y=m,ymin=ymin,ymax=ymax))
}
mean_summary <- function(x) {
   m <- mean(x)
   return(c(y=m))
}
```

# Load and process

```{r, message=FALSE}
# Osc
freq <- "20"

# Init
results <- NULL

# high w
exp <- "100"
name <- paste("../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(2.48e-08, nrow(df))
results <- rbind(df, results)

# Ref (Z=0)
exp <- "96"
name <- paste("../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(8.25e-09, nrow(df))
results <- rbind(df, results)

# low w
exp <- "99"
name <- paste("../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(1.65e-09, nrow(df))
results <- rbind(df, results)

# Process
results %>% 
  filter(n_spikes_ref > 0) %>% 
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  mutate(V_osc = V_osc - V_osc_ref) -> results

rm(df, exp, freq, name)
```

# As

```{r, fig.width=1.9, fig.height=1.5}
results %>% 
  # filter(rank == 0) %>% 
  ggplot(aes(x=As, y=n_spikes_pop, color=factor(weight))) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  # filter(rank == 0) %>% 
  ggplot(aes(x=As, y=variances_pop, color=factor(weight))) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  # filter(rank == 0) %>% 
  ggplot(aes(x=As, y=errors_pop, color=factor(weight))) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

# V_osc/V_comp

```{r, fig.width=1.9, fig.height=1.3}
results %>% 
  filter(V_osc/V_comp <= 3) %>% 
  group_by(N, weight) %>% 
  mutate(norm_errors = errors/max(errors)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_errors, group=interaction(N, weight), color=factor(weight*1e9))) +
  geom_line(size=0.3, alpha=0.4) + 
  theme_classic() +
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  # scale_colour_brewer(palette="Greys", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```