### 2018-04-04 voltagebudget/analysis/paper_figures.Rmd
Added control exps
Hetero vis still in progress

---
title: "Exp1 - Opt w_max and {A,phi)"
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-mad(x)
   ymax <- m+mad(x)
   return(c(y=m, ymin=ymin, ymax=ymax))
}

mean_summary <- function(x) {
   m <- mean(x)
   return(c(y=m))
}
sample_n_groups = function(tbl, size, replace = FALSE, weight = NULL) {
  # regroup when done
  grps = tbl %>% groups %>% lapply(as.character) %>% unlist
  # check length of groups non-zero
  keep = tbl %>% summarise() %>% ungroup() %>% sample_n(size, replace, weight)
  # keep only selected groups, regroup because joins change count.
  # regrouping may be unnecessary but joins do something funky to grouping variable
  tbl %>% right_join(keep, by=grps) %>% group_by_(.dots = grps)
}
```

# Read and process all data

```{r, message=FALSE}
# --------------------------------------------------------------------
# Example voltage timecourses
fig0 <- read_csv("vm_comp_timecourse.csv")

# --------------------------------------------------------------------
fig1 <- NULL

# --------------------------
# INITIAL
exp <- "96"
freqs <- c(20, 30)

for(freq in freqs){
  name <- paste("../data/exp", exp, "_d-2e-3_f", as.character(freq), ".csv", sep="")
  df <- read_csv(paste(name, sep=","))
  df %>%
    filter(n_spikes > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig1 <- rbind(fig1, df)
}

# --------------------------
# SLOW
exp <- "94"
freqs <- c(6, 8, 10, 12, 14, 16, 18)

for(freq in freqs){
  name <- paste("../data/exp", exp, "_d-2e-3_f", as.character(freq), ".csv", sep="")
  df <- read_csv(paste(name, sep=","))
  df %>%
    filter(n_spikes > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig1 <- rbind(fig1, df)
}

# --------------------------
# FAST
exp <- "95"
freqs <- c(22, 24, 26, 28)#, 32, 34 ,36, 38)
for(freq in freqs){
  name <- paste("../data/exp", exp, "_d-2e-3_f", as.character(freq), ".csv", sep="")
  df <- read_csv(paste(name, sep=","))
  df %>%
    filter(n_spikes > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig1 <- rbind(fig1, df)
}

# Process
fig1 %>% 
  filter(n_spikes_ref > 0) %>% 
  # Add delta_spikes
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  # Re-norm V_osc
  mutate(V_osc = V_osc - V_osc_ref) -> fig1

# Summarize over As 
fig1 %>% 
  group_by(As, freq) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig2

# --------------------------------------------------------------------
# Data
algs <- c("max", "left", "uni", "cc")
alg_names <- c("max", "left", "uniform", "coincidence")
exps <- c(82, 83, 84, 85)

fig3 <- NULL
for (n in 1:4){
  df <- read_csv(paste("../data/exp", as.character(exps[n]), "_", algs[n], ".csv", sep=""))
  df$alg <- rep(alg_names[n], nrow(df))

  fig3 <- rbind(fig3, df)
}

# Spikes
# Load ref spikes
ref <- read_csv("../data/stim1.csv")
ref$set <- rep("initial", nrow(ref))

# Load exp spikes
exps <- c(86, 87, 88, 89)
fig4 <- NULL
for (n in 1:4){
  df <- read_csv(paste("../data/exp", as.character(exps[n]), "_", algs[n], "_c0.1_spikes.csv", sep=""))
  df$set <- rep("synchronized", nrow(df))

  # Add ref
  df <- rbind(df, ref)

  df$alg <- rep(alg_names[n], nrow(df))
  fig4 <- rbind(fig4, df)
}

# Shift-add cartoon
fig5 <- read_csv("vm_ref_timecourse.csv")
fig6 <- read_csv("vm_shift_timecourse.csv")
fig7 <- read_csv("vm_add_timecourse.csv")

# --------------------------------------------------------------------
# Load noise exp data...
# Osc
freq <- "20"

# sigma scan
exp <- "97"

sigmas <- c("1.0e-9", "2.0e-9", "3.0e-9", "4.0e-9", "5.0e-9")

fig8 <- NULL
for(sigma in sigmas){
  name <- paste("../data/exp", exp, "_sigma", sigma, ".csv", sep="")
  df <- read_csv(paste(name, sep=","))

  df$sigma <- rep(as.numeric(sigma), nrow(df))
  fig8 <- rbind(fig8, df)
}

exp <- "96"
name <- paste("../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$sigma <- rep(0, nrow(df))
fig8 <- rbind(df, fig8)

# Process
fig8 %>% 
  filter(n_spikes_ref > 0) %>% 
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  mutate(V_osc = V_osc - V_osc_ref) -> fig8

# Summarize over As 
fig8 %>% 
  group_by(As, sigma) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig9

# --------------------------------------------------------------------
# Load weight exps
# Osc
freq <- "20"

# Init
fig10 <- NULL

# high w
exp <- "100"
name <- paste("../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(2.48e-08, nrow(df))
fig10 <- rbind(df, fig10)

# Ref (Z=0)
exp <- "96"
name <- paste("../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(8.25e-09, nrow(df))
fig10 <- rbind(df, fig10)

# low w
exp <- "99"
name <- paste("../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(1.65e-09, nrow(df))
fig10 <- rbind(df, fig10)

# Process
fig10 %>% 
  filter(n_spikes_ref > 0) %>% 
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  mutate(V_osc = V_osc - V_osc_ref) -> fig10

# Summarize over As 
fig10 %>% 
  group_by(As, weight) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig11

# --------------------------------------------------------------------
# Hetero cells
# Osc
freq <- "20"

# Init
fig12 <- NULL

# high w
exp <- "101"
name <- paste("../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(2.48e-08, nrow(df))
fig12 <- rbind(df, fig12)

# Process
fig12 %>% 
  filter(n_spikes_ref > 0) %>% 
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  mutate(V_osc = V_osc - V_osc_ref) -> fig12

# Summarize over As 
fig12 %>% 
  group_by(As, weight) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig13

# --------------------------------------------------------------------
rm(df, exp, freq, freqs, name, ref, n, sigmas, sigma, exps, alg_names, algs)
```

# Shift-add cartoon

```{r, fig.width=1.2, fig.height=0.5}
t_start <- 0.2
t_stop <- 0.3

fig6 %>% 
  filter(t > t_start, t < t_stop) %>%
  ggplot(aes(x=t, ymin=-70e-3*1e3, ymax=osc*1e3)) +
  geom_ribbon(fill="red", alpha=0.5) +
  geom_ribbon(aes(x=t, ymin=osc*1e3, ymax=comp*1e3), fill="purple", alpha=0.5) +
  geom_hline(yintercept=-55e-3*1e3, color="black", linetype="dotted", size=0.5) +
  geom_hline(yintercept=-70e-3*1e3, color="black", linetype="dotted", size=0.5) +
  labs(x="", y="") +
  theme_classic() +
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      # axis.title.x=element_blank(),
      # axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank()) -> p1

fig7 %>% 
  filter(t > t_start, t < t_stop) %>%
  ggplot(aes(x=t, ymin=-70e-3*1e3, ymax=osc*1e3)) +
  geom_ribbon(fill="red", alpha=0.5) +
  geom_ribbon(aes(x=t, ymin=osc*1e3, ymax=comp*1e3), fill="purple", alpha=0.5) +
  geom_hline(yintercept=-55e-3*1e3, color="black", linetype="dotted", size=0.5) +
  geom_hline(yintercept=-70e-3*1e3, color="black", linetype="dotted", size=0.5) +
  labs(x="", y="") +
  theme_classic() +
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      # axis.title.x=element_blank(),
      # axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank()) -> p2

plot_grid <- rbind(c(1, 1, 1, 2, 2, 2))
grid.arrange(p1, p2, layout_matrix=plot_grid)
rm(p1, p2, plot_grid)
```

# Budget intro

```{r, fig.width=1.5, fig.height=1.3}
fig0 %>% 
  filter(t > 0.15, t < 0.35) %>%
  ggplot(aes(x=t, ymin=-70e-3*1e3, ymax=vm*1e3)) +
  geom_ribbon(fill="blue", alpha=0.5) +
  geom_hline(yintercept=-50e-3*1e3, color="black", linetype="dotted", size=0.5) +
  geom_hline(yintercept=-70e-3*1e3, color="black", linetype="dotted", size=0.5) +
  labs(x="", y=expression(V[m] (mvolt))) +
  theme_classic() -> p1

fig0 %>% 
  filter(t > 0.15, t < 0.35) %>%
  ggplot(aes(x=t, ymin=-70e-3*1e3, ymax=osc*1e3)) +
  geom_ribbon(fill="red", alpha=0.5) +
  geom_ribbon(aes(x=t, ymin=osc*1e3, ymax=comp*1e3), fill="purple", alpha=0.5) +
  geom_hline(yintercept=-50e-3*1e3, color="black", linetype="dotted", size=0.5) +
  geom_hline(yintercept=-70e-3*1e3, color="black", linetype="dotted", size=0.5) +
  geom_rect(aes(xmin=0.23652 - 8e-3, xmax=0.23652 + 2e-3 + 0.03333333 + 8e-3, ymin=-70e-3*1e3, ymax=-50e-3*1e3),  fill=NA, color="black", alpha=0.01, size=.2) +
  # geom_hline(yintercept=-70e-3, color="black", size=1.5) +
  # geom_rect(aes(xmin=0.23652 - 2e-3, xmax=0.23652, ymin=-70e-3, ymax=-50e-3), color="grey", alpha=0.5) +
  # geom_rect(aes(xmin=0.23652 + 2e-3, xmax=0.23652 + 2e-3 + 0.03333333, ymin=-70e-3, ymax=-50e-3), fill=NA, color="grey", alpha=0.1) +
  labs(x="Time (s)", y="") +
  theme_classic() -> p2

fig0 %>% 
  filter(t > 0.23652 - 8e-3, t < 0.23652 + 2e-3 + 0.03333333 + 8e-3) %>%
  ggplot(aes(x=t, ymin=-70e-3*1e3, ymax=osc*1e3)) +
  geom_ribbon(fill="red", alpha=0.5) +
  geom_ribbon(aes(x=t, ymin=osc*1e3, ymax=comp*1e3), fill="purple", alpha=0.5) +
  # geom_hline(yintercept=-50e-3, color="black", linetype="dotted", size=0.5) +
  # geom_hline(yintercept=-70e-3, color="black", linetype="dotted", size=0.5) +
  geom_rect(aes(xmin=0.23652 - 2e-3, xmax=0.23652, ymin=-70e-3*1e3, ymax=-50e-3*1e3), fill="grey", alpha=0.01) +
  geom_rect(aes(xmin=0.23652 + 2e-3, xmax=0.23652 + 2e-3 + 0.03333333, ymin=-70e-3*1e3, ymax=-50e-3*1e3),  fill="grey", alpha=0.01) +
  geom_rect(aes(xmin=0.23652 - 8e-3, xmax=0.23652 + 2e-3 + 0.03333333 + 8e-3, ymin=-70e-3*1e3, ymax=-50e-3*1e3),  fill=NA, color="black", alpha=0.01, size=.2) +
  labs(x="", y="") +
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      # axis.title.x=element_blank(),
      # axis.title.y=element_blank(),
      legend.position="none",
      panel.background=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.background=element_blank()) -> p3

plot_grid <- rbind(c(1 , 1, NA), c(2, 2, 3))
grid.arrange(p1, p2, p3, layout_matrix=plot_grid)
rm(p1, p2, p3, plot_grid)
```

```{r, fig.width=1, fig.height=0.8}
fig1 %>%
  filter(rank == 0, freq == 20) %>%
  ggplot(aes(x=As*1e9, y=V_osc*1e3)) +
  geom_point(size=0.1, alpha=0.6, color="black") +
  labs(x="A (namp)", y=expression(V[o] (mvolt))) +
  theme_classic() 
```
# Trade-off

```{r, fig.width=1.5, fig.height=1.9}
fig2 %>% 
  filter(freq == 20) %>% 
  ggplot(aes(x=V_osc/V_b, y=errors_pop*1e3, color=V_osc/V_b)) +
  geom_point(size=0.05, alpha=1) + 
  labs(x=expression(V[o]/V[b]), y="Avg. error (ms)") +
  # lims(y=c(0, 10)) +
  scale_x_continuous(breaks=c(0, 0.5, 1.0), limits = c(0, 1)) +
  scale_color_gradient(low="darkgrey", high="red") +
  theme_classic() +
  theme(
    legend.position="none") -> p1

fig2 %>% 
  filter(freq == 20) %>% 
  ggplot(aes(x=V_osc/V_b, y=variances_pop*1e3, color=V_osc/V_b)) +
  geom_point(size=0.05, alpha=1) + 
  labs(x=expression(V[o]/V[b]), y="Pop. variance (ms)") +
  scale_color_gradient(low="darkgrey", high="red") +
  scale_x_continuous(breaks=c(0, 0.5, 1.0), limits = c(0, 1)) +
  theme_classic() +
  theme(
    legend.position="none") -> p2

fig2 %>% 
  filter(freq == 20) %>% 
  ggplot(aes(x=errors_pop*1e3, y=variances_pop*1e3, color=V_osc/V_b)) +
  geom_point(size=0.05, alpha=.6) +
  scale_color_gradient(low="darkgrey", high="red", name=expression(V[o]/V[b])) +
  # scale_color_gradient(low="darkgrey", high="red", name="A (namp)") +
  geom_smooth(color="red", method=lm, alpha=0.2, size=.15, se=FALSE, linetype='solid') +
  labs(x="Avg. error (ms)", y="Pop. variance (ms)") +
  theme_classic() +
  theme(
    legend.position="right", 
    legend.background = element_rect(colour ="black"),
    legend.key.size = unit(.4, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8)) -> p3

# plot_grid <- rbind(
#   c(NA, NA, NA, NA, 4, 4,   4),
#   c(NA, NA, NA, NA, 4, 4,   4),
#   c(1,  1,  1,  1, 2, 2,   2),
#   c(1,  1,  1,  1, 2, 2,   2),
#   c(1,  1,  1,  1, 3, 3,   3),
#   c(1,  1,  1,  1, 3, 3,   3))#,
  # c(1,  1,  1,  NA, NA, NA))

plot_grid <- rbind(
  c(NA),
  c(1, 2),
  c(1, 2),
  c(3, 3),
  c(3, 3),
  c(3, 3))

grid.arrange(p1, p2, p3, layout_matrix=plot_grid)
rm(p1, p2, p3, plot_grid)
```

# Bound 1 - V_osc/V_comp (f = 20)

```{r, fig.width=2.7, fig.height=2.2}
fig1 %>% 
  group_by(N) %>% 
  filter(V_osc/V_comp <= 3) %>% 
  filter(freq == 20) %>% 
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  ggplot(aes(x=V_osc/V_comp, y=errors*1e3, group=N)) +
  geom_line(size=0.5, alpha=0.2, color="grey") + 
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  theme_classic() +
  geom_line(data=filter(fig2, freq==20, V_osc/V_comp <= 3), 
            aes(x=V_osc/V_comp, y=errors*1e3), 
            color="black", size=0.8) +
  labs(x=expression(V[o]/V[c]), y="Error (ms)") -> p1

fig1 %>% 
  filter(V_osc/V_comp <= 3) %>% 
  filter(freq == 20) %>% 
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  ggplot(aes(x=V_osc/V_comp, y=variances*1e3, group=N)) +
  geom_line(size=0.5, alpha=0.2, color="grey") + 
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  theme_classic() +
  geom_line(data=filter(fig1, freq==20, V_osc/V_comp <= 3), 
            aes(x=V_osc/V_comp, y=variances_pop*1e3), 
            color="black", size=0.7, alpha=1) +
  labs(x=expression(V[o]/V[c]), y="Variance (ms)") -> p2

max_N <- max(fig1$N)

fig1 %>% 
  filter(freq==20) %>% 
  filter(V_osc/V_comp <= 3) %>% 
  filter(N %in% sample(1:max_N, 24, replace = FALSE)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=errors*1e3, group=N)) +
  geom_line(size=.5, alpha=0.9, color="grey") + 
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  facet_wrap(~rank, nrow=3) +
  theme_classic() + 
  scale_x_continuous(breaks=c(0, 1, 2)) +
  scale_y_continuous(breaks=c(0, 10)) +
  theme(
    strip.text.x = element_text(size=0),
    strip.background = element_blank(),
    # axis.line=element_blank(),
    # axis.text.x=element_blank(),
    # axis.text.y=element_blank(),
    # axis.ticks=element_blank(),s
    # axis.title.x=element_blank(),
    # axis.title.y=element_blank(),
    panel.background=element_blank(),
    # panel.border=element_blank(),
    # panel.grid.major=element_blank(),
    # panel.grid.minor=element_blank(),
    plot.background=element_blank()) +
  labs(x=expression(V[o]/V[c]), y="Error (ms)") -> p3

plot_grid <- rbind(
  c(1, 1, 1, 2, 2), 
  c(1, 1, 1, 2, 2), 
  c(1, 1, 1, 2, 2),
  c(1, 1, 1, NA, NA),
  c(3, 3, 3, 3, 3),
  c(3, 3, 3, 3, 3),
  c(3, 3, 3, 3, 3))
  # c(3, 3, 3, 4, 4),
  # c(3, 3, 3, 4, 4),
  # c(3, 3, 3, 4, 4))
grid.arrange(p1, p2, p3, layout_matrix=plot_grid)
rm(p1, p2, p3, plot_grid, max_N)
```

# Bound 2 - n_spikes < (0, 1)

```{r, fig.width=2.6, fig.height=3.2}
max_N <- max(fig1$N)

fig1 %>% 
  filter(freq == 20) %>% 
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  ggplot(aes(x=V_osc/V_b, y=variances*1e3, group=N)) +
  geom_line(size=0.5, alpha=0.2, color="grey") + 
  theme_classic() +
  geom_line(aes(x=V_osc/V_b, y=variances_pop*1e3), 
            color="black", size=0.7, alpha=1) +
  labs(x=expression(V[o]/V[b]), y="Variance (ms)") -> p1

fig1 %>% 
  filter(freq == 20) %>% 
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  ggplot(aes(x=V_osc/V_b, y=variances*1e3, group=N, color=factor(delta_spikes))) +
  geom_line(size=0.4, alpha=0.6) + 
  scale_colour_brewer(
    palette = "Reds", guide=guide_legend(title="Spike count", nrow=3)) +
  labs(x=expression(V[o]/V[b]), y="Variance (ms)") +
  theme_classic() + theme(legend.position="none") -> p2

# -
M = min(fig1$variances_pop)
fig1 %>% 
  filter(freq == 20) %>% 
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  ggplot(aes(x=variances*1e3, fill=factor(delta_spikes))) +
  geom_density(alpha=0.6, size=0, color="black") +
  scale_fill_brewer(palette = "Reds", name="Spike\ncount") +
  geom_vline(xintercept = M*1e3, color="black", linetype="dotted") +
  labs(x="Variance (ms)", y="Density (AU)") +
  theme_classic() + 
  theme(legend.position = c(1.25, 1.55), 
        legend.background = element_rect(colour ="black"),
        legend.key.size = unit(.5, "cm"),
        legend.title = element_text(size=10),
        legend.text = element_text(size=10)) -> p3

fig1 %>% 
  filter(freq==20) %>% 
  filter(V_osc/V_comp <= 3) %>% 
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  ggplot(aes(y=V_osc/V_comp, x=factor(delta_spikes), group=delta_spikes)) +
  geom_jitter(aes(color=factor(delta_spikes)), size=0.01, alpha=0.1, width=0.2) +
  geom_boxplot(outlier.colour = NA, alpha=0.9, width=0.2) +
  scale_colour_brewer(palette = "Reds", name="Spike count") +
  geom_hline(yintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_hline(yintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  theme_classic() +
  theme(legend.position = "none") +
  labs(y=expression(V[o]/V[c]), x="Spike count") + 
  coord_flip()  -> p4

fig1 %>%
  filter(freq==20) %>%
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  ungroup() %>% 
  group_by(N) %>% 
  sample_n_groups(8) %>% 
  ggplot(aes(x=V_osc/V_b, y=variances*1e3, color=factor(delta_spikes))) +
  geom_point(size=1, alpha=1) +
  facet_wrap(~N, nrow=8) +
  geom_line(aes(x=V_osc/V_b, y=variances_pop*1e3),
            color="black", size=0.7, alpha=1) +
  scale_colour_brewer(palette = "Reds", name="Spike count") +
  scale_x_continuous(breaks=c(0, 0.5, 1.0), limits = c(0, 1)) +
  scale_y_continuous(breaks=c(6, 12)) +
  labs(x=expression(V[o]/V[b]), y="Variance (ms)") +
  theme_classic() + theme(
    # legend.position="bottom",
    strip.text.x = element_text(size=0),
    strip.background = element_blank(),
    legend.position="none",
    legend.key.size = unit(.5, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8),
    # axis.line=element_blank(),
    # axis.text.x=element_blank(),
    # axis.text.y=element_blank(),
    # axis.ticks=element_blank(),
    # axis.title.x=element_blank(),
    # axis.title.y=element_blank(),
    panel.background=element_blank(),
    # panel.border=element_blank(),
    # panel.grid.major=element_blank(),
    # panel.grid.minor=element_blank(),
    plot.background=element_blank()) -> p5

fig1 %>%
  filter(freq==20) %>%
  filter(V_osc/V_comp <= 3) %>%
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  sample_n_groups(8) %>% 
  ggplot(aes(x=V_osc/V_b, y=errors*1e3, color=factor(delta_spikes))) +
  geom_point(size=1, alpha=1) +
  facet_wrap(~N, nrow=8) +
  # geom_line(aes(x=V_osc/V_b, y=errors_pop*1e3),
  #           color="black", size=0.7, alpha=1) +
  scale_colour_brewer(palette = "Reds", name="Spike count") +
  scale_x_continuous(breaks=c(0, 0.5, 1), limits = c(0, 1)) +
  scale_y_continuous(breaks=c(2, 8)) +
  labs(x=expression(V[o]/V[b]), y="Error (ms)") +
  theme_classic() + theme(
    # legend.position="bottom",
    strip.text.x = element_text(size=0),
    strip.background = element_blank(),
    legend.position="none",
    legend.key.size = unit(.5, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8),
    # axis.line=element_blank(),
    # axis.text.x=element_blank(),
    # axis.text.y=element_blank(),
    # axis.ticks=element_blank(),
    # axis.title.x=element_blank(),
    # axis.title.y=element_blank(),
    panel.background=element_blank(),
    # panel.border=element_blank(),
    # panel.grid.major=element_blank(),
    # panel.grid.minor=element_blank(),
    plot.background=element_blank()) -> p6

plot_grid <- rbind(
  c(1, 1, 1, NA, 5, 5, 6, 6),
  c(1, 1, 1, NA, 5, 5, 6, 6),
  c(1, 1, 1, NA, 5, 5, 6, 6),
  c(2, 2, 2, NA, 5, 5, 6, 6),
  c(2, 2, 2, NA, 5, 5, 6, 6),
  c(2, 2, 2, NA, 5, 5, 6, 6),
  c(3, 3, 3, NA, 5, 5, 6, 6),
  c(3, 3, 3, NA, 5, 5, 6, 6),
  c(3, 3, 3, NA, 5, 5, 6, 6),
  c(4, 4, 4, NA, NA, NA, NA, NA),
  c(4, 4, 4, NA, NA, NA, NA, NA),
  c(4, 4, 4, NA, NA, NA, NA, NA))

grid.arrange(p1, p2, p3, p4, p5, p6, layout_matrix=plot_grid)
rm(p1, p2, p3, p4, plot_grid, max_N, M)
```

# Control exps
## Freqs, noise, weight

```{r, fig.width=2.1, fig.height=2.4}
# --------------------------------------------------------------
# Freq
fig2 %>% 
  filter(V_osc/V_comp < 3) %>% 
  filter(freq <= 30) %>% 
  group_by(N, freq) %>% 
  mutate(norm_errors_pop = errors_pop/max(errors_pop)) %>% 
  mutate(norm_variances_pop = variances_pop/max(variances_pop)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_errors_pop, color=freq, group=interaction(N, freq))) +
  geom_line(size=0.3, alpha=1) + 
  labs(x=expression(V[o]/V[b]), y="Norm. error") +
  scale_color_gradient2(low="gold", high="darkgreen", mid="grey", midpoint=20) +
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  theme_classic() +
  theme(legend.position="none", 
    legend.key.size = unit(.4, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8)) -> p1

fig2 %>% 
  filter(V_osc/V_comp < 3) %>% 
  filter(freq <= 30) %>% 
  group_by(N, freq) %>% 
  mutate(norm_errors_pop = errors_pop/max(errors_pop)) %>% 
  mutate(norm_variances_pop = variances_pop/max(variances_pop)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_variances_pop, color=freq, group=interaction(N, freq))) +
  geom_line(size=0.3, alpha=1) + 
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  labs(x=expression(V[o]/V[b]), y="Norm. variance") +
  scale_color_gradient2(low="gold", high="darkgreen", mid="grey", midpoint=20, name="Freq. (Hz)") +
  theme_classic() +
  # lims(y=c(5, 10)) +
  theme(legend.position="right", 
    legend.background = element_rect(colour ="black"),
    legend.key.size = unit(.4, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8)) -> p2

# --------------------------------------------------------------
# Noise
fig9 %>% 
  filter(V_osc/V_comp < 3) %>% 
  group_by(N, sigma) %>% 
  mutate(norm_errors_pop = errors_pop/max(errors_pop)) %>% 
  mutate(norm_variances_pop = variances_pop/max(variances_pop)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_errors_pop, color=10e-3*(sigma/10e-9) * 1e3/2, group=interaction(N, sigma))) +
  geom_line(size=0.3, alpha=1) + 
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  labs(x=expression(V[o]/V[b]), y="Norm. error") +
  scale_color_gradient(low="grey", high="orange") +
  theme_classic() +
  theme(legend.position="none", 
    legend.key.size = unit(.4, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8)) -> p3

fig9 %>% 
  filter(V_osc/V_comp < 3) %>% 
  group_by(N, sigma) %>% 
  mutate(norm_errors_pop = errors_pop/max(errors_pop)) %>% 
  mutate(norm_variances_pop = variances_pop/max(variances_pop)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_variances_pop, color=10e-3*(sigma/10e-9) * 1e3 /2, group=interaction(N, sigma))) +
  geom_line(size=0.3, alpha=1) + 
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  labs(x=expression(V[o]/V[b]), y="Norm. variance") +
  scale_color_gradient(low="grey", high="orange", name="Noise (mV)") +
  theme_classic() +
  # lims(y=c(5, 10)) +
  theme(legend.position="right", 
    legend.background = element_rect(colour ="black"),
    legend.key.size = unit(.4, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8)) -> p4

# --------------------------------------------------------------
# Weight
fig11 %>% 
  filter(V_osc/V_comp < 3) %>% 
  group_by(N, weight) %>% 
  mutate(norm_errors_pop = errors_pop/max(errors_pop)) %>% 
  mutate(norm_variances_pop = variances_pop/max(variances_pop)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_errors_pop, color=weight*1e9, group=interaction(N, weight))) +
  geom_line(size=0.3, alpha=1) + 
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  labs(x=expression(V[o]/V[b]), y="Norm. error") +
  scale_color_gradient2(
    low="cornflowerblue", high="darkblue", mid="grey", midpoint=8.25e-09*1e9) +
  theme_classic() +
  theme(legend.position="none", 
    legend.key.size = unit(.4, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8)) -> p5

fig11 %>% 
  filter(V_osc/V_comp < 3) %>% 
  group_by(N, weight) %>% 
  mutate(norm_errors_pop = errors_pop/max(errors_pop)) %>% 
  mutate(norm_variances_pop = variances_pop/max(variances_pop)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_variances_pop, color=weight*1e9, group=interaction(N, weight))) +
  geom_line(size=0.3, alpha=1) + 
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.5, color="black", linetype="dashed") +
  labs(x=expression(V[o]/V[b]), y="Norm. variance") +
  scale_color_gradient2(
    low="cornflowerblue", high="darkblue", mid="grey", midpoint=8.25e-09*1e9, name="Weight (nS)") +
  theme_classic() +
  # lims(y=c(5, 10)) +
  theme(legend.position="right", 
    legend.background = element_rect(colour ="black"),
    legend.key.size = unit(.4, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8)) -> p6

plot_grid <- rbind(
  c(1, 1, 1, 2, 2, 2, 2, 2),
  c(3, 3, 3, 4, 4, 4, 4, 4),
  c(5, 5, 5, 6, 6, 6, 6, 6))

grid.arrange(p1, p2, p3, p4, p5, p6, layout_matrix=plot_grid)
rm(p1, p2, p3, p4, plot_grid)
```
## Heterocells
```{r, fig.width=1.1, fig.height=1.5}
fig12 %>%
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  filter(V_osc/V_comp <= 3) %>% 
  mutate(norm_errors = errors/max(errors)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_errors, group=N)) +
  geom_line(size=0.5, alpha=0.2, color="darkgrey") + 
  geom_line(data=filter(fig13, V_osc/V_comp <= 3), 
            aes(x=V_osc/V_comp, y=errors/max(errors)), 
            color="black", size=0.8) +
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.4, color="black", linetype="dashed") +
  theme_classic() +
  lims(x=c(0, 3)) +
  labs(x=expression(V[o]/V[b]), y="Norm. error") -> p1

fig12 %>%
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  filter(V_osc/V_comp <= 3) %>% 
  mutate(norm_variances = variances/max(variances)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_variances, group=N)) +
  geom_line(size=0.5, alpha=0.2, color="darkgrey") + 
  geom_line(data=filter(fig13, V_osc/V_comp <= 3), 
            aes(x=V_osc/V_comp, y=variances/max(variances)), 
            color="black", size=0.8) +
  geom_vline(xintercept = 1, alpha=0.7, color="black", linetype="dashed") +
  geom_vline(xintercept = 0.5, alpha=0.4, color="black", linetype="dashed") +
  theme_classic() +
  lims(x=c(0, 3)) +
  labs(x=expression(V[o]/V[b]), y="Norm. variance") -> p2

plot_grid <- rbind(
  c(1),
  c(2))

grid.arrange(p1, p2, layout_matrix=plot_grid)
rm(p1, p2, plot_grid)
```

# Optimal

```{r, fig.width=3, fig.height=1.4}
fig4 %>% 
  filter(alg != "coincidence") %>%
  ggplot(aes(x=(ts - min(ts))*1e3, y=ns, color=set)) +
  geom_point(size=0.7, alpha=1) +
  facet_grid(alg~.) +
  scale_color_manual(values=c("black", "darkgrey"), name="") +
  theme_classic() +
  labs(x="Time (ms)", y="Neuron") +
  theme(
    legend.position="top", 
    legend.key.size = unit(.5, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0)) -> p1

fig3 %>% 
  filter(alg != "coincidence") %>%
  ggplot(aes(x=errors*1e3, y=obs_variances*1e3, linetype=alg)) +
  geom_line(size=0.8, alpha=1, color="darkgrey") +
  scale_linetype_manual(values=c("dashed", "solid", "dotted"), name="Algorithm") +
  theme_classic() +
  labs(x="Error (ms)", y="Variance (ms)") -> p2

plot_grid <- rbind(
  c(1, 1, 1, NA, NA, NA, NA),
  c(1, 1, 1, 2, 2, 2, 2),
  c(1, 1, 1, 2, 2, 2, 2),
  c(1, 1, 1, 2, 2, 2, 2))

grid.arrange(p1, p2, layout_matrix=plot_grid)
rm(p1, p2, plot_grid)
```