### 2018-12-02 voltagebudget/analysis/fixed_run_time_analysis.Rmd
Update figs

---
title: "Sweep A - general experiment analysis."
output: html_notebook
---

```{r, message=FALSE}
library(tidyverse)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}
```

# Load and process

```{r, message=FALSE}
results <- NULL

# --------------------------"
exp <- "102"
freqs <- c(10, 20, 30)
run_times <-  c("0.25", "0.5", "1", "1.5", "2", "2.5", "3.0")
for(freq in freqs){
  for(t in run_times) {
    # exp102_f10_t1.5.csv
    name <- paste("../data/exp", exp, "_f", freq, "_t", t, ".csv", sep="")
    df <- read_csv(paste(name, sep=","))
    df %>%
      filter(n_spikes > 0) -> df

    df$freq <- rep(as.numeric(freq), nrow(df))
    df$run_time <- rep(as.numeric(t), nrow(df))
    results <- rbind(results, df)  
  }
}

# --------------------------
# Process
results %>% 
  # Add delta_spikes
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  # Re-norm V_osc
  mutate(V_osc = V_osc - V_osc_ref) -> results

# --------------------------
# Est avg pop responses
results %>% 
  group_by(freq, run_time) %>% 
  summarise(
    N = 0,
    V_b = mean(V_b),
    V_osc = mean(V_osc),
    n_spikes = mean(n_spikes),
    n_spikes_pop = mean(n_spikes_pop),
    delta_spikes = mean(delta_spikes),
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    V_comp = mean(V_comp), 
    errors = mean(errors), 
    variances_pop = mean(variances_pop),
    variances = mean(variances)) -> avg_results

rm(df, exp, freq, freqs, t, run_times, name)
```

# Population

```{r, fig.width=1.5, fig.height=0.9}
avg_results %>%
  ggplot(aes(x=run_time, y=(variances_pop/n_spikes_pop)*1e3, color=freq, group=interaction(freq, run_time))) +
  geom_point(size=0.3, alpha=0.9) +
  labs(x="Run time (s)", y="Avg. variance (ms)") +
  theme_classic() +
  scale_color_gradient2(low="gold", high="darkgreen", mid="grey", midpoint=20, name="Freq. (Hz)") +
  theme(legend.position="right", 
    legend.key.size = unit(.4, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank()) 
```

# Individual cells

```{r, fig.width=2.5, fig.height=2.7}
results %>%
  filter(As < 0.05e-9) %>% 
  ggplot(aes(x=As, y=(variances/n_spikes)*1e3, color=run_time, group=interaction(As, run_time))) +
  # geom_point(size=0.05, alpha=0.1) +
  geom_boxplot(outlier.size = NA) +
  labs(x="A (namp)", y="Variance (ms)") +
  theme_classic() +
  facet_grid(freq~.) +
  scale_color_gradient(low="blue", high="darkgreen") +
  lims(y=c(0, 20)) +
  theme(legend.position="bottom", 
    legend.key.size = unit(.4, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank()) 
```