### 2019-03-27 voltagebudget/analysis/optimal_sweep_A_analysis.Rmd
Added opt_sweep_A notebook

---
title: "Optimal coordination - general experiment analysis."
output: html_notebook
---

```{r, message=FALSE}
library(ggplot2)
library(magrittr)
library(readr)
library(dplyr)
library(gridExtra)
library(readr)
library(scales)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}
```

# Load and process

```{r, message=FALSE}
# Data
exp <- "103"
freqs <- c("4", "8", "12", "20", "24", "30")#, "40", "60")

results <- NULL
for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  # df %>%
    # filter(n_spikes_pop > 0) -> df

  df$freq <- rep(freq, nrow(df))
  results <- rbind(results, df)
}

results$freq <- factor(results$freq, levels=freqs)
rm(df, exp, freq, freqs, name)
```

```{r, fig.width=1, fig.height=3}
ms <- 1e3
results %>% 
  filter(errors_pop > 0) %>% 
  ggplot(aes(x=errors_pop*ms, y=variances_pop*ms)) +
  geom_line(size=0.5) + 
  geom_line(mapping=aes(x=errors_opt*ms, y=variances_opt*ms), color="#e41a1c", alpha=0.8, size=.5) +
  scale_x_continuous(trans='log10') +
  annotation_logticks(sides = "b") +
  facet_grid(freq~.) +
  labs(x="Pop. error (ms)", y="Pop. variance (ms)") +
  theme_classic() +
  theme(strip.text.y = element_text(angle = 0), 
        panel.grid.minor = element_blank(),
         strip.background = element_blank())
```