### 2019-05-19 voltagebudget/analysis/paper_figures_redraft.Rmd
Update fig redraft

---
title: "Paper figures"
output: html_notebook
---

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(readr)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-mad(x)
   ymax <- m+mad(x)
   return(c(y=m, ymin=ymin, ymax=ymax))
}

mean_summary <- function(x) {
   m <- mean(x)
   return(c(y=m))
}
sample_n_groups = function(tbl, size, replace = FALSE, weight = NULL) {
  # regroup when done
  grps = tbl %>% groups %>% lapply(as.character) %>% unlist
  # check length of groups non-zero
  keep = tbl %>% summarise() %>% ungroup() %>% sample_n(size, replace, weight)
  # keep only selected groups, regroup because joins change count.
  # regrouping may be unnecessary but joins do something funky to grouping variable
  tbl %>% right_join(keep, by=grps) %>% group_by_(.dots = grps)
}

ms <- 1000
mV <- 1000
```

# Read and process all data

__NOTE__: to generate the needed 'vm_*.csv' files (which are voltage timecourses used in various figs) you have to 'run all' cells in the following:
 1. `../ipynb/figures_shit_add.ipynb`
 2. `../ipynb/figures_heterogenous.ipynb`
 3. `../ipynb/figures_SFN.ipynb`

```{r, message=FALSE}
# --------------------------------------------------------------------
# Example voltage timecourses
fig0 <- read_csv("vm_comp_timecourse.csv")

# --------------------------------------------------------------------
fig1 <- NULL

# --------------------------
# INITIAL
exp <- "96"
freqs <- c(20, 30)

for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", as.character(freq), ".csv", sep="")
  df <- read_csv(paste(name, sep=","))
  df %>%
    filter(n_spikes > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig1 <- rbind(fig1, df)
}

# --------------------------
# SLOW
exp <- "94"
freqs <- c(6, 8, 10, 12, 14, 16, 18)

for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", as.character(freq), ".csv", sep="")
  df <- read_csv(paste(name, sep=","))
  df %>%
    filter(n_spikes > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig1 <- rbind(fig1, df)
}

# --------------------------
# FAST
exp <- "95"
freqs <- c(22, 24, 26, 28)#, 32, 34 ,36, 38)
for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", as.character(freq), ".csv", sep="")
  df <- read_csv(paste(name, sep=","))
  df %>%
    filter(n_spikes > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig1 <- rbind(fig1, df)
}

# Process
fig1 %>% 
  filter(n_spikes_ref > 0) %>% 
  # Add delta_spikes
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  # Re-norm V_osc
  mutate(V_osc = V_osc - V_osc_ref) -> fig1

# Summarize over As 
fig1 %>% 
  group_by(N, freq) %>% 
  mutate(errors=normalize(errors),
            variances=normalize(variances)) -> fig1

fig1 %>% 
  group_by(As, freq) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_free = mean(V_free), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    n_spikes_pop = mean(n_spikes), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig2

# --------------------------------------------------------------------
# Data
algs <- c("max", "left", "uni", "cc")
alg_names <- c("max", "left", "uniform", "coincidence")
exps <- c(82, 83, 84, 85)

fig3 <- NULL
for (n in 1:4){
  df <- read_csv(paste("../../data/exp", as.character(exps[n]), "_", algs[n], ".csv", sep=""))
  df$alg <- rep(alg_names[n], nrow(df))

  fig3 <- rbind(fig3, df)
}

# Spikes
# Load ref spikes
ref <- read_csv("../../data/stim1.csv")
ref$set <- rep("initial", nrow(ref))

# Load exp spikes
exps <- c(86, 87, 88, 89)
fig4 <- NULL
for (n in 1:4){
  df <- read_csv(paste("../../data/exp", as.character(exps[n]), "_", algs[n], "_c0.1_spikes.csv", sep=""))
  df$set <- rep("synchronized", nrow(df))

  # Add ref
  df <- rbind(df, ref)

  df$alg <- rep(alg_names[n], nrow(df))
  fig4 <- rbind(fig4, df)
}

# Shift-add cartoon
fig5 <- read_csv("vm_ref_timecourse.csv")
fig6 <- read_csv("vm_shift_timecourse.csv")
fig7 <- read_csv("vm_add_timecourse.csv")

# --------------------------------------------------------------------
# Load noise exp data...
# Osc
freq <- "20"

# sigma scan
exp <- "97"

sigmas <- c("1.0e-9", "2.0e-9", "3.0e-9", "4.0e-9", "5.0e-9")

fig8 <- NULL
for(sigma in sigmas){
  name <- paste("../../data/exp", exp, "_sigma", sigma, ".csv", sep="")
  df <- read_csv(paste(name, sep=","))

  df$sigma <- rep(as.numeric(sigma), nrow(df))
  fig8 <- rbind(fig8, df)
}

exp <- "96"
name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$sigma <- rep(0, nrow(df))
fig8 <- rbind(df, fig8)

# Process
fig8 %>% 
  filter(n_spikes_ref > 0) %>% 
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  mutate(V_osc = V_osc - V_osc_ref) -> fig8

# Summarize over As 
fig8 %>% 
  group_by(As, sigma) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    n_spikes_pop = mean(n_spikes), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig9

# --------------------------------------------------------------------
# Load weight exps
# Osc
freq <- "20"

# Init
fig10 <- NULL

# high w
exp <- "100"
name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(2.48e-08, nrow(df))
fig10 <- rbind(df, fig10)

# Ref (Z=0)
exp <- "96"
name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(8.25e-09, nrow(df))
fig10 <- rbind(df, fig10)

# low w
exp <- "99"
name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(1.65e-09, nrow(df))
fig10 <- rbind(df, fig10)

# Process
fig10 %>% 
  filter(n_spikes_ref > 0) %>% 
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  mutate(V_osc = V_osc - V_osc_ref) -> fig10

# Summarize over As 
fig10 %>% 
  group_by(As, weight) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    n_spikes_pop = mean(n_spikes), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig11

# --------------------------------------------------------------------
# Hetero cells
# Osc
freq <- "20"

# Init
fig12 <- NULL

# high w
exp <- "101"
name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(2.48e-08, nrow(df))
fig12 <- rbind(df, fig12)

# Process
fig12 %>% 
  filter(n_spikes_ref > 0) %>% 
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  mutate(V_osc = V_osc - V_osc_ref) -> fig12

# Summarize over As 
fig12 %>% 
  group_by(As, weight) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    n_spikes_pop = mean(n_spikes), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig13

# Load exmample timecourses for hetero
fig14 <- read_csv("vm_hetero_timecourse.csv")

# -------------------------------------------------------------------
# Optimal versus oscillation

# Data
exp <- "104"
freqs <- c("4", "8", "12", "20", "24", "30")#, "40", "60")

fig15 <- NULL
for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  # df %>%
    # filter(n_spikes_pop > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig15 <- rbind(fig15, df)
}

fig15$freq <- factor(fig15$freq, levels=freqs)

exp <- "105"
freqs <- c("4", "8", "12", "20", "24", "30")#, "40", "60")

fig16 <- NULL
for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  # df %>%
    # filter(n_spikes_pop > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig16 <- rbind(fig16, df)
}

fig16$freq <- factor(fig16$freq, levels=freqs)

fig15$alg <- rep("max", nrow(fig15))
fig16$alg <- rep("uniform", nrow(fig16))
fig17 <- rbind(fig15, fig16)
# --------------------------------------------------------------------
rm(df, exp, freq, freqs, name, ref, n, sigmas, sigma, exps, alg_names, algs)
```

# The trade off

```{r, fig.width=1.7, fig.height=2.8}
# Var
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  ggplot(aes(x=V_osc*mV, variances_pop*ms, group=freq)) +
  geom_point(size=.2, alpha=0.5) +
  xlab(expression(paste(V[o], " (mV)"))) +
  ylab("Variance (ms)") +
  theme_classic() -> p1
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  ggplot(aes(x=V_osc/V_comp, variances_pop*ms, group=freq)) +
  geom_point(size=.2, alpha=0.5) +
  geom_vline(xintercept = 0.27, alpha=1, color="darkslategray3", linetype="solid") +
  xlab(expression(V[o]/V[c])) +
  ylab("Variance (ms)") +
  theme_classic() -> p2

# Error
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  ggplot(aes(x=V_osc*mV, errors_pop*ms, group=freq)) +
  geom_point(size=.2, alpha=0.5) +
  xlab(expression(paste(V[o], " (mV)"))) +
  ylab("Error (ms)") +
  theme_classic() -> p3
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  ggplot(aes(x=V_osc/V_comp, errors_pop*ms, group=freq)) +
  geom_point(size=.2, alpha=0.5) +
  geom_vline(xintercept = 0.27, alpha=1, color="darkslategray3", linetype="solid") +
  xlab(expression(V[o]/V[c])) +
  ylab("Error (ms)") +
  theme_classic() -> p4

tmp <- fig2
tmp$state <- rep("pathological", nrow(tmp))
m <- tmp$V_osc/tmp$V_comp < 0.27
tmp$state[m] <- "healthy"

tmp %>% 
  filter(V_osc/V_comp <= 5) %>%
  ggplot(aes(x=variances_pop*ms, errors_pop*ms, group=freq, color=state)) +
  geom_point(size=.2, alpha=0.9) +
  xlab("Variance (ms)") +
  ylab("Error (ms)") +
  scale_color_manual("", values=c("darkslategray3", "darkslategray")) +
  theme_classic() -> p5

plot_grid <- rbind(
  c(NA, NA, NA, NA),
  c(1, 1, 3, 3), 
  c(1, 1, 3, 3),
  c(1, 1, 3, 3),
  c(1, 1, 3, 3),
  c(NA, NA, NA, NA),
  c(2, 2, 4, 4),
  c(2, 2, 4, 4),
  c(2, 2, 4, 4),
  c(2, 2, 4, 4),
  c(NA, NA, NA, NA),
  c(5, 5, 5, 5),
  c(5, 5, 5, 5),
  c(5, 5, 5, 5),
  c(5, 5, 5, 5),
  c(5, 5, 5, 5),
  c(5, 5, 5, 5))

grid.arrange(p1, p2, p3, p4, p5, layout_matrix=plot_grid)
```

By looking at the ratio we identified a narrow region (V_o/V_c ~< 0.27) where the trade-off is approx linear across freq, and that accounts for approx. half the total change in both variance and error. We hypothesize that this narrow region can be considered a healthy range for modulation as it is maximially sensitive and linear -- representing the best possible trade-off irregardless of freq.

Sync efficiency is strongly dependent on other osc input. Strong synapses need strong oscillations. Weak synapses need weak oscillations. To stay in healthy range, oscillation strength should be tuned on a compartment basis, either dendritically or somatically or both.

# Vc and sync

```{r, fig.width=1.4, fig.height=1}
fig2 %>% 
  # filter(V_osc/V_comp <= ) %>%
  ggplot(aes(x=V_comp*mV, variances_pop*ms, color=V_osc*mV)) +
  geom_point(size=.2, alpha=0.5) +
  xlab(expression(paste(V[c], " (mV)"))) +
  ylab("Variance (ms)") +
  scale_color_gradient(expression(paste(V[o], " (mV)")), low="goldenrod1", high="firebrick3") +
  theme_classic() +
  theme(legend.key.width=unit(0.3,"cm"),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8))
```

While the ratio determines the overall trade-off (above), again we see that about half the sync effect comes from weak oscillations (< 1 mV, orange) applied to strong inputs (i.e., V_c > 6 mV). We interpret this to mean that the optimal target for oscillatory modulation are strong synapses. However this interpretation is complicated by the fact that the weaker Vc, the more sync is possible--though this sync is very high error.

# Freq. dependence
```{r, fig.width=1, fig.height=6}
fig2 %>% 
  filter(V_osc/V_comp <= 0.8) %>%
  ggplot(aes(x=V_osc/V_comp, y=variances_pop*ms)) +
  geom_point(size=.2, alpha=0.5) +
  facet_grid(freq~.) +
  xlab(expression(V[o]/V[c])) +
  ylab("Variance (ms)") +
  # geom_rect(
  #   mapping=aes(xmin=0.15, xmax=0.4, ymin=min(variances_pop*ms), ymax=max(variances_pop*ms)),
  #   color="darkslategray4", fill="darkslategray4", alpha=0.02) +
  geom_vline(xintercept = 0.27, alpha=1, color="darkslategray3", linetype="solid") +
  theme_classic() +
  theme(strip.text.y = element_text(angle=0), strip.background = element_blank())
```

Fo far we've focused on error-variance relationships common between frequencies. There are also clearly frequency dependent effects

...

Faster oscillations offer less sync overall, and therefore have a more limited var-error range than slower oscillations. If we use the curveture of the variance curve to fine tune the healthy ratio / frequency, we see that...

TODO: get max second deriv. Use that to find the change point?

# Controls - Ws, noise, cell type
TODO

# MINST exp (BIND)
TODO

# Optimal derivation
```{r, fig.width=1.3, fig.height=2}
min_t <- min(fig4$ts)

fig4 %>% 
  filter(alg != "coincidence", alg != "left") %>% 
  group_by(set, alg) %>% 
  mutate(ts = ts - min_t) -> tmp

tmp %>% 
  filter(set == "synchronized") %>%
  ggplot(aes(x=ts*1e3, y=ns, color=alg)) +
  geom_point(size=2, alpha=1, shape=4) +
  geom_point(
    data=filter(tmp, set == "initial"),
    mapping=aes(x=ts*1e3, y=ns), color="black", shape=1, alpha=1, size=1.2) +
  facet_grid(alg~.) +
  scale_color_manual("Algorithm:", values=c("darkslategray4", "darkseagreen3")) +
  theme_classic() +
  labs(x="Time (ms)", y="Neuron") +
  theme(
    legend.position="none", 
    panel.border = element_rect(fill = NA, colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0)) -> p1

fig3 %>% 
  filter(alg != "coincidence", alg != "left") %>%
  ggplot(aes(x=errors*1e3, y=obs_variances*1e3, color=alg)) +
  geom_line(size=0.6, alpha=0.99) +
  scale_color_manual("Algorithm:", values=c("darkslategray4", "darkseagreen3")) +
  theme_classic() +
  coord_flip() +
  labs(x="Time (ms)", y="Neuron") +
  theme(legend.position="bottom",
    legend.key.size = unit(.2, "cm"),
    legend.title = element_text(size=10),
    legend.text = element_text(size=10),
    legend.background = element_rect(colour ="black")) +
  labs(x="Error (ms)", y="Variance (ms)") -> p2

plot_grid <- rbind(
  c(1, 1, 1, 1, 1),
  c(1, 1, 1, 1, 1),
  c(2, 2, 2, 2, NA),
  c(2, 2, 2, 2, NA))

grid.arrange(p1, p2, layout_matrix=plot_grid)
rm(p1, p2, plot_grid)
```

# Optimal vs. oscillation

```{r, fig.width=1.3, fig.height=1.2}
ms <- 1e3

fig17 %>% 
  filter(errors_pop > 0, freq == "20") %>% 
  ggplot(aes(x=variances_pop*ms, y=errors_pop*ms)) +
  geom_point(size=0.2, alpha=0.99) + 
  geom_point(mapping=aes(x=variances_opt*ms, y=errors_opt*ms, color=alg), 
             alpha=0.99, size=.2) +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l") +
  scale_color_manual("Algorithm:", values=c("darkslategray4", "darkseagreen3")) +
  labs(x="Variance (ms)", y="Error (ms)") +
  theme_classic() +
  theme(legend.position="bottom",
    legend.key.size = unit(.2, "cm"),
    legend.title = element_text(size=10),
    legend.text = element_text(size=10),
    legend.background = element_rect(colour ="black"),
    strip.text.y = element_text(angle = 0), 
    panel.grid.minor = element_blank(),
    legend.key.width=unit(0.1,"cm"),
    strip.background = element_blank()) 
```