### 2019-05-21 voltagebudget/analysis/exp107_analysis.Rmd

---
title: "Sweep A - general experiment analysis."
output: html_notebook
---

```{r, message=FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}

mV <- 1000
ms <- 1000
```

# Load and process

```{r, message=FALSE}
exp <- "107"
V_ls <- c("-70e-3", "-72e-3", "-74e-3", "-76e-3", "-78e-3", "-80e-3")

results <- NULL
for(V_l in V_ls){
  name <- paste("../../data/exp", exp, "_V_l", V_l, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  df %>%
    filter(n_spikes > 0) -> df

  df$V_l <- rep(V_l, nrow(df))
  results <- rbind(results, df)
}
results$V_l <- factor(results$V_l, levels=V_ls)

results %>% 
  filter(n_spikes_ref > 0) %>% 
  # Add delta_spikes
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  # Re-norm V_osc
  mutate(V_osc = V_osc - V_osc_ref) -> results

# -
rm(df, exp, V_l, V_ls, name)
```

# Overall distributions

```{r, fig.width=2.75, fig.height=2}
results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=variances_pop*ms, fill=V_l)) +
  geom_histogram(bins=20, alpha=1) + 
  xlab("Variance") +
  theme_classic() -> p1

results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=errors_pop*ms, fill=V_l)) +
  geom_histogram(bins=20) + 
  xlab("Error") +
  theme_classic() -> p2

results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=V_osc*mV, fill=V_l)) +
  geom_histogram(bins=20) +
  xlab("Vc (volt)") +
  theme_classic() -> p3

results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=V_comp*mV, fill=V_l)) +
  geom_histogram(bins=20) +
  xlab("Vo (volt)") +
  theme_classic() -> p4

grid.arrange(p1, p2, p3, p4, nrow=2)
rm(p1, p2, p3, p4)
```

# As

```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=As*1e9, y=V_osc*mV, color=V_l)) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=V_osc*mV, y=n_spikes_pop - n_spikes_ref, color=V_l)) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=V_osc * mV, y=variances_pop*ms, color=V_l)) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```
```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=V_osc/V_comp, y=variances_pop*ms, color=V_l)) +
  geom_vline(xintercept = 0.27, alpha=1, color="darkslategray3", linetype="solid") +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.6, fig.height=1.6}
results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=V_osc * mV, y=errors_pop*ms, color=V_l)) +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.6, fig.height=1.6}
results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=V_osc/V_comp, y=errors_pop*ms, color=V_l)) +
  geom_vline(xintercept = 0.27, alpha=1, color="darkslategray3", linetype="solid") +
  geom_line() + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```

```{r, fig.width=1.5, fig.height=1.5}
results %>% 
  filter(N == 1, V_osc/V_comp < 5) %>% 
  ggplot(aes(x=variances_pop*ms, y=errors_pop*ms, color=V_l)) +
  geom_point(size=.05) + 
  theme_classic() +
  scale_colour_brewer(palette="Greens", guide=guide_legend(nrow=3)) +
  theme(legend.position="bottom") 
```