### 2019-05-21 voltagebudget/analysis/paper_figures_redraft.Rmd
WIP

---
title: "Paper figures"
output: html_notebook
---

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(readr)
library(gridExtra)
library(readr)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-mad(x)
   ymax <- m+mad(x)
   return(c(y=m, ymin=ymin, ymax=ymax))
}

mean_summary <- function(x) {
   m <- mean(x)
   return(c(y=m))
}
sample_n_groups = function(tbl, size, replace = FALSE, weight = NULL) {
  # regroup when done
  grps = tbl %>% groups %>% lapply(as.character) %>% unlist
  # check length of groups non-zero
  keep = tbl %>% summarise() %>% ungroup() %>% sample_n(size, replace, weight)
  # keep only selected groups, regroup because joins change count.
  # regrouping may be unnecessary but joins do something funky to grouping variable
  tbl %>% right_join(keep, by=grps) %>% group_by_(.dots = grps)
}

ms <- 1000
mV <- 1000
```

# Read and process all data

__NOTE__: to generate the needed 'vm_*.csv' files (which are voltage timecourses used in various figs) you have to 'run all' cells in the following:
 1. `../ipynb/figures_shit_add.ipynb`
 2. `../ipynb/figures_heterogenous.ipynb`
 3. `../ipynb/figures_SFN.ipynb`

```{r, message=FALSE}
# --------------------------------------------------------------------
# Example voltage timecourses
fig0 <- read_csv("vm_comp_timecourse.csv")

# --------------------------------------------------------------------
fig1 <- NULL

# --------------------------
# INITIAL
exp <- "96"
freqs <- c(20, 30)

for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", as.character(freq), ".csv", sep="")
  df <- read_csv(paste(name, sep=","))
  df %>%
    filter(n_spikes > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig1 <- rbind(fig1, df)
}

# --------------------------
# SLOW
exp <- "94"
freqs <- c(6, 8, 10, 12, 14, 16, 18)

for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", as.character(freq), ".csv", sep="")
  df <- read_csv(paste(name, sep=","))
  df %>%
    filter(n_spikes > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig1 <- rbind(fig1, df)
}

# --------------------------
# FAST
exp <- "95"
freqs <- c(22, 24, 26, 28)#, 32, 34 ,36, 38)
for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", as.character(freq), ".csv", sep="")
  df <- read_csv(paste(name, sep=","))
  df %>%
    filter(n_spikes > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig1 <- rbind(fig1, df)
}

# Process
fig1 %>% 
  filter(n_spikes_ref > 0) %>% 
  # Add delta_spikes
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  # Re-norm V_osc
  mutate(V_osc = V_osc - V_osc_ref) -> fig1

# Summarize over As 
fig1 %>% 
  group_by(N, freq) %>% 
  mutate(errors=normalize(errors),
            variances=normalize(variances)) -> fig1

fig1 %>% 
  group_by(As, freq) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_free = mean(V_free), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    n_spikes_pop = mean(n_spikes), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig2

# Ratio ests
fig2_ratios <- NULL
freqs <- unique(fig2$freq)
for (f in freqs) {
  tmp <- filter(fig2, f == freq, V_osc/V_comp < 5)
  pnts <- find_point(normalize(tmp$errors_pop), tmp$V_osc/tmp$V_comp, 0.5)  
  row <- c(pnts, f)
  fig2_ratios <- rbind(fig2_ratios, row)
}
fig2_ratios <- as.data.frame(fig2_ratios, stringsAsFactors=FALSE)
colnames(fig2_ratios) <- c("error_50", "ratio", "freq")

# --------------------------------------------------------------------
# Data
algs <- c("max", "left", "uni", "cc")
alg_names <- c("max", "left", "uniform", "coincidence")
exps <- c(82, 83, 84, 85)

fig3 <- NULL
for (n in 1:4){
  df <- read_csv(paste("../../data/exp", as.character(exps[n]), "_", algs[n], ".csv", sep=""))
  df$alg <- rep(alg_names[n], nrow(df))

  fig3 <- rbind(fig3, df)
}

# Spikes
# Load ref spikes
ref <- read_csv("../../data/stim1.csv")
ref$set <- rep("initial", nrow(ref))

# Load exp spikes
exps <- c(86, 87, 88, 89)
fig4 <- NULL
for (n in 1:4){
  df <- read_csv(paste("../../data/exp", as.character(exps[n]), "_", algs[n], "_c0.1_spikes.csv", sep=""))
  df$set <- rep("synchronized", nrow(df))

  # Add ref
  df <- rbind(df, ref)

  df$alg <- rep(alg_names[n], nrow(df))
  fig4 <- rbind(fig4, df)
}

# Shift-add cartoon
fig5 <- read_csv("vm_ref_timecourse.csv")
fig6 <- read_csv("vm_shift_timecourse.csv")
fig7 <- read_csv("vm_add_timecourse.csv")

# --------------------------------------------------------------------
# Load noise exp data...
# Osc
freq <- "20"

# sigma scan
exp <- "97"

sigmas <- c("1.0e-9", "2.0e-9", "3.0e-9", "4.0e-9", "5.0e-9")

fig8 <- NULL
for(sigma in sigmas){
  name <- paste("../../data/exp", exp, "_sigma", sigma, ".csv", sep="")
  df <- read_csv(paste(name, sep=","))

  df$sigma <- rep(as.numeric(sigma), nrow(df))
  fig8 <- rbind(fig8, df)
}

exp <- "96"
name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$sigma <- rep(0, nrow(df))
fig8 <- rbind(df, fig8)

# Process
fig8 %>% 
  filter(n_spikes_ref > 0) %>% 
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  mutate(V_osc = V_osc - V_osc_ref) -> fig8

# Summarize over As 
fig8 %>% 
  group_by(As, sigma) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    n_spikes_pop = mean(n_spikes), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig9

# --------------------------------------------------------------------
# Hetero cells
# Osc
freq <- "20"

# Init
fig12 <- NULL

# high w
exp <- "101"
name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
df <- read_csv(paste(name, sep=","))
df$weight <- rep(2.48e-08, nrow(df))
fig12 <- rbind(df, fig12)

# Process
fig12 %>% 
  filter(n_spikes_ref > 0) %>% 
  mutate(delta_spikes = n_spikes - n_spikes_ref) %>% 
  mutate(V_osc = V_osc - V_osc_ref) -> fig12

# Summarize over As 
fig12 %>% 
  group_by(As, weight) %>% 
  summarise(
    N = 0,
    delta_spikes = mean(delta_spikes), 
    V_b = mean(V_b), 
    V_osc = mean(V_osc), 
    V_comp = mean(V_comp), 
    V_comp_ref = mean(V_comp_ref), 
    n_spikes_pop = mean(n_spikes), 
    errors = mean(errors), 
    errors_pop = mean(errors_pop), 
    variances_pop = mean(variances_pop), 
    variances = mean(variances)) -> fig13

# Load exmample timecourses for hetero
fig14 <- read_csv("vm_hetero_timecourse.csv")

# -------------------------------------------------------------------
# Optimal versus oscillation

# Data
exp <- "104"
freqs <- c("4", "8", "12", "20", "24", "30")#, "40", "60")

fig15 <- NULL
for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  # df %>%
    # filter(n_spikes_pop > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig15 <- rbind(fig15, df)
}

fig15$freq <- factor(fig15$freq, levels=freqs)

exp <- "105"
freqs <- c("4", "8", "12", "20", "24", "30")#, "40", "60")

fig16 <- NULL
for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  # df %>%
    # filter(n_spikes_pop > 0) -> df

  df$freq <- rep(freq, nrow(df))
  fig16 <- rbind(fig16, df)
}

fig16$freq <- factor(fig16$freq, levels=freqs)

fig15$alg <- rep("max", nrow(fig15))
fig16$alg <- rep("uniform", nrow(fig16))
fig17 <- rbind(fig15, fig16)
# --------------------------------------------------------------------
rm(df, exp, freq, freqs, name, ref, n, sigmas, sigma, exps, alg_names, algs)
```

# The trade off
```{r, fig.width=2, fig.height=5}
# By freq
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  ggplot(aes(x=V_osc*mV, errors_pop*ms, group=freq, color=freq)) +
  geom_line(size=.5, alpha=0.8) +
  scale_colour_continuous("Freq.", low="lightsalmon1", high="orange4") +
  xlab(expression(paste(V[o], " (mV)"))) +
  ylab("Error (ms)") +
  theme_classic() +
  theme(legend.position="none") -> p1
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  ggplot(aes(x=V_osc*mV, variances_pop*ms, group=freq, color=freq)) +
  geom_line(size=.5, alpha=0.8) +
  scale_colour_continuous("Freq.", low="lightsalmon1", high="orange4") +
  xlab(expression(paste(V[o], " (mV)"))) +
  ylab("Variance (ms)") +
  theme_classic() +
  theme(
    legend.position=c(0.45, 1.5),
    legend.direction="horizontal",    
    legend.key.size = unit(.5, "cm"),
    legend.key.height=unit(0.3,"cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8),
    legend.background = element_rect(colour ="black")) -> p2

# Norm
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  group_by(freq, N) %>%
  mutate(errors_pop = normalize(errors_pop)) %>%
  mutate(variances_pop = normalize(variances_pop)) %>%
  ggplot(aes(x=V_osc/V_comp, errors_pop, group=freq)) +
  geom_line(size=.5, alpha=0.5) +
  geom_hline(yintercept = 0.5, alpha=1, color="grey", linetype="dashed") +
  annotate("text", x = 4, y = 0.65, label = expression(paste("E"[50]))) +
  xlab(expression(V[o]/V[c])) +
  ylab("Norm. error") +
  theme_classic() -> p3
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  group_by(freq, N) %>%
  mutate(errors_pop = normalize(errors_pop)) %>%
  mutate(variances_pop = normalize(variances_pop)) %>%
  ggplot(aes(x=V_osc/V_comp, variances_pop, group=freq)) +
  geom_line(size=.5, alpha=0.5) +
  xlab(expression(V[o]/V[c])) +
  ylab("Norm. variance") +
  theme_classic() -> p4

# Budget
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  group_by(freq, N) %>%
  mutate(errors_pop = normalize(errors_pop)) %>%
  mutate(variances_pop = normalize(variances_pop)) %>%
  ggplot(aes(x=V_osc/V_comp, errors_pop, group=freq)) +
  geom_line(size=.5, alpha=0.5) +

  # Line
  geom_vline(xintercept = 0.2858, alpha=0.8, color="darkslategray3", linetype="solid") +
  xlab(expression(V[o]/V[c])) +
  ylab("Norm. error") +
  theme_classic() -> p5
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  group_by(freq, N) %>%
  mutate(errors_pop = normalize(errors_pop)) %>%
  mutate(variances_pop = normalize(variances_pop)) %>%
  ggplot(aes(x=V_osc/V_comp, variances_pop, group=freq)) +
  geom_line(size=.5, alpha=0.5) +

  # Line
  geom_vline(xintercept = 0.2858, alpha=0.8, color="darkslategray3", linetype="solid") +
  xlab(expression(V[o]/V[c])) +
  ylab("Norm. variance") +
  theme_classic() -> p6

tmp <- fig2
tmp$state <- rep("pathological", nrow(tmp))
m <- tmp$V_osc/tmp$V_comp < 0.2858
tmp$state[m] <- "healthy"

tmp %>% 
  filter(V_osc/V_comp <= 5) %>%
  group_by(freq, N) %>%
  mutate(errors_pop = normalize(errors_pop)) %>%
  mutate(variances_pop = normalize(variances_pop)) %>%
  ggplot(aes(x=variances_pop, errors_pop, group=freq, color=state)) +
  geom_point(size=.8, alpha=0.8) +
  xlab("Norm. variance") +
  ylab("Norm. error") +
  scale_color_manual("", values=c("darkslategray3", "darkslategray")) +
  theme_classic() -> p7

print(summary(fig2_ratios$ratio))
fig2_ratios %>% 
  ggplot(aes(x=freq, y=ratio)) +
  geom_point(color="darkslategray3", size=1.6) +
  xlab("Freq (Hz)") +
  geom_hline(yintercept = mean(fig2_ratios$ratio), alpha=0.8, color="darkslategray3", linetype="solid") +
  ylab(expression(paste(V[o]/V[c], " at E"[50]))) +
  theme_classic() +
  ylim(0, 1) -> p8

plot_grid <- rbind(
  c(NA, NA, NA, NA),
  c(NA, NA, NA, NA),
  c(1, 1, 2, 2), 
  c(1, 1, 2, 2),
  c(1, 1, 2, 2),
  c(1, 1, 2, 2),
  c(NA, NA, NA, NA),
  c(3, 3, 4, 4),
  c(3, 3, 4, 4),
  c(3, 3, 4, 4),
  c(3, 3, 4, 4),
  c(NA, NA, NA, NA),
  c(8, 8, 8, NA),
  c(8, 8, 8, NA),
  c(8, 8, 8, NA),
  c(8, 8, 8, NA),
  c(NA, NA, NA, NA),
  c(5, 5, 6, 6),
  c(5, 5, 6, 6),
  c(5, 5, 6, 6),
  c(5, 5, 6, 6),
  c(NA, NA, NA, NA),
  c(7, 7, 7, 7),
  c(7, 7, 7, 7),
  c(7, 7, 7, 7),
  c(7, 7, 7, 7),
  c(7, 7, 7, 7),
  c(7, 7, 7, 7),
  c(7, 7, 7, 7))

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, layout_matrix=plot_grid)
```

Error below 0.5 implies the result contains a majority of the reference. Beyond 0.5, the modulator dominates spiking. This means 0.5 serves as natural boundry to seperate oscillations that modulate other activity, and oscillations that activity. The ratio at 0.5 error is V_o/V_c = 0.27. This point turns out to several interesting features. It is where the trade-off is approx linear across freq, and where the trade-off is most stable.

We hypothesize that this narrow region of budget ratio can be considered a healthy range for modulation as it, let's comp dom, is maximially sensitive (has the highest dirvative) and is approx. linear is maximmally consistent--representing the best possible trade-off irregardless of freq.

Sync efficiency is strongly dependent on other osc input. Strong synapses need strong oscillations. Weak synapses need weak oscillations. To stay in healthy range, oscillation strength should be tuned on a compartment basis, either dendritically or somatically or both.

Our analysis suggests a simple biological scaling rule for oscillatory modulation. This is useful both in analyzing for patholigical acitivty, but also in desgining optogenetic and other stimulation experiments. You should scale your oscillatory manipulation be no more the 0.2858 times as strong as the average membrane potential observed in vivo, without stimulation.

If your observaing a excitatory interneuron, with weak sparse inputs, a good level of oscillatory stimulation will be much lower then if, for example, you were working with a layer 5 pyrimidal cell recieving strong thalmic drive.

# Vc and sync
```{r, fig.width=1.4, fig.height=0.8}
fig2 %>% 
  filter(V_osc/V_comp <= 5) %>%
  group_by(freq, N) %>% 
  mutate(errors_pop = normalize(errors_pop)) %>%
  mutate(variances_pop = normalize(variances_pop)) %>%
  ggplot(aes(x=V_comp*mV, variances_pop, color=V_osc*mV)) +
  geom_point(size=.2, alpha=0.5) +
  xlab(expression(paste(V[c], " (mV)"))) +
  ylab("Norm. variance") +
  scale_color_gradient(expression(paste(V[o], " (mV)")), low="darkgrey", high="firebrick3") +
  theme_classic() +
  theme(legend.key.height=unit(0.3,"cm"),
        legend.key.width=unit(0.3,"cm"),
        legend.background = element_rect(colour ="black"),
        legend.title=element_text(size=8),
        legend.text=element_text(size=8))
```

While the ratio determines the overall trade-off (above), again we see that about half the sync effect comes from weak oscillations (< 1 mV, orange) applied to strong inputs (i.e., V_c > 6 mV). We interpret this to mean that the optimal target for oscillatory modulation are strong synapses. However this interpretation is complicated by the fact that the weaker Vc, the more sync is possible--though this sync is very high error. That is, if very high syncrony is needed very high error unavoidably the cost--a cost we can predict based on measurable membrane potential Vc.

A simple biological response to this is to have (weak, <1 mV) oscillatory modulation selectivly target strong synapses.

# Controls - Ws, noise, cell type
```{r, fig.width=2.1, fig.height=1.5}
# --------------------------------------------------
fig14 %>% 
  ggplot(aes(x=t, y=vm*1e3, group=cell)) +
  geom_line(size=0.2, color="black") +
  facet_grid(cell~.) +
  theme_classic() +
  labs(x="Time (s)", y="V (mV)") +
  theme(
    strip.text = element_blank(),
    strip.background = element_blank(),
    legend.position="none",
    axis.line=element_blank(),
    axis.text.x=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks=element_blank(),
    axis.title.x=element_blank(),
    axis.title.y=element_blank(),
    panel.background=element_blank(),
    panel.border=element_blank(),
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank(),
    plot.background=element_blank()) -> p1

# Cell type
fig12 %>%
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  filter(V_osc/V_comp <= 5) %>% 
  mutate(norm_errors = normalize(errors_pop)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_errors, group=N)) +
  geom_point(size=.1, alpha=0.4) +
  geom_vline(xintercept = 0.2858, alpha=1, color="darkslategray3", linetype="solid") +
  labs(x=expression(V[o]/V[c]), y="Norm. error", title="Cell type") +
  theme_classic() +
  theme(title = element_text(size=10)) +
  lims(x=c(0, 5)) -> p2

fig12 %>%
  group_by(N) %>%
  mutate(V_free_initial = round(max(V_free), 3)) %>%
  filter(V_free_initial > 0) %>%
  filter(V_osc/V_comp <= 5) %>% 
  mutate(norm_variance = normalize(variances_pop)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_variance, group=N)) +
  geom_point(size=.1, alpha=0.4) +
  geom_vline(xintercept = 0.2858, alpha=1, color="darkslategray3", linetype="solid") +
  theme_classic() +
  lims(x=c(0, 5)) +
  labs(x=expression(V[o]/V[c]), y="Norm. variance") -> p3

# --------------------------------------------------------------
# Noise
fig9 %>% 
  filter(V_osc/V_comp < 5) %>% 
  group_by(N, sigma) %>% 
  mutate(norm_errors_pop = normalize(errors_pop)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_errors_pop, 
             color=10e-3*(sigma/10e-9) * 1e3/2, group=interaction(N, sigma))) +
  geom_point(size=.1, alpha=0.4) +
  geom_vline(xintercept = 0.2858, alpha=1, color="darkslategray3", linetype="solid") +
  labs(x=expression(V[o]/V[c]), y="Norm. error") +
  scale_color_gradient("Noise\n(mV)", low="grey", high="black") +
  theme_classic() +
  # theme(legend.position="top", 
  theme(
    # legend.position = c(.2, .9), 
    # legend.direction = "horizontal",
    legend.position="top", 
    legend.margin=margin(t = 0, b=-.25, unit='cm'),
    legend.key.size = unit(.3, "cm"),
    legend.key.height=unit(0.15,"cm"),
    legend.key.width=unit(0.3,"cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=4)) -> p4

fig9 %>% 
  filter(V_osc/V_comp < 5) %>% 
  group_by(N, sigma) %>% 
  mutate(norm_variances_pop = normalize(variances_pop)) %>% 
  ggplot(aes(x=V_osc/V_comp, y=norm_variances_pop, 
             color=10e-3*(sigma/10e-9) * 1e3/2, group=interaction(N, sigma))) +
  geom_point(size=.1, alpha=0.4) +
  geom_vline(xintercept = 0.2858, alpha=1, color="darkslategray3", linetype="solid") +
  labs(x=expression(V[o]/V[c]), y="Norm. variance") +
  scale_color_gradient(low="grey", high="black") +
  theme_classic() +
  theme(legend.position="none") -> p5

plot_grid <- rbind(
  c(1, 1, 2, 2, 2, 2, 4, 4, 4, 4),
  c(1, 1, 2, 2, 2, 2, 4, 4, 4, 4),
  c(1, 1, 2, 2, 2, 2, 4, 4, 4, 4),
  c(1, 1, 2, 2, 2, 2, 4, 4, 4, 4),
  c(1, 1, 2, 2, 2, 2, 4, 4, 4, 4),
  c(1, 1, 3, 3, 3, 3, 5, 5, 5, 5),
  c(1, 1, 3, 3, 3, 3, 5, 5, 5, 5),
  c(1, 1, 3, 3, 3, 3, 5, 5, 5, 5),
  c(1, 1, 3, 3, 3, 3, 5, 5, 5, 5))

grid.arrange(p1, p2, p3, p4, p5, layout_matrix=plot_grid)
```

# MINST exp (BIND)
TODO

# Optimal derivation
```{r, fig.width=1.3, fig.height=2}
min_t <- min(fig4$ts)

fig4 %>% 
  filter(alg != "coincidence", alg != "left") %>% 
  group_by(set, alg) %>% 
  mutate(ts = ts - min_t) -> tmp

tmp %>% 
  filter(set == "synchronized") %>%
  ggplot(aes(x=ts*1e3, y=ns, color=alg)) +
  geom_point(size=2, alpha=1, shape=4) +
  geom_point(
    data=filter(tmp, set == "initial"),
    mapping=aes(x=ts*1e3, y=ns), color="black", shape=1, alpha=1, size=1.2) +
  facet_grid(alg~.) +
  scale_color_manual("Algorithm:", values=c("darkslategray4", "darkseagreen3")) +
  theme_classic() +
  labs(x="Time (ms)", y="Neuron") +
  theme(
    legend.position="none", 
    panel.border = element_rect(fill = NA, colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0)) -> p1

fig3 %>% 
  filter(alg != "coincidence", alg != "left") %>%
  ggplot(aes(x=errors*1e3, y=obs_variances*1e3, color=alg)) +
  geom_line(size=0.6, alpha=0.99) +
  scale_color_manual("Algorithm:", values=c("darkslategray4", "darkseagreen3")) +
  theme_classic() +
  coord_flip() +
  labs(x="Time (ms)", y="Neuron") +
  theme(legend.position="bottom",
    legend.key.size = unit(.2, "cm"),
    legend.title = element_text(size=10),
    legend.text = element_text(size=10),
    legend.background = element_rect(colour ="black")) +
  labs(x="Error (ms)", y="Variance (ms)") -> p2

plot_grid <- rbind(
  c(1, 1, 1, 1, 1),
  c(1, 1, 1, 1, 1),
  c(2, 2, 2, 2, NA),
  c(2, 2, 2, 2, NA))

grid.arrange(p1, p2, layout_matrix=plot_grid)
rm(p1, p2, plot_grid)
```

# Optimal vs. oscillation

```{r, fig.width=1.1, fig.height=3}
ms <- 1e3

fig17 %>% 
  filter(errors_pop > 0, freq %in% c("8","12","20","30")) %>%
  ggplot(aes(x=variances_pop*ms, y=errors_pop*ms)) +
  geom_point(size=0.2, alpha=0.99) + 
  geom_point(mapping=aes(x=variances_opt*ms, y=errors_opt*ms, color=alg), 
             alpha=0.99, size=.2) +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l") +
  facet_grid(freq~.) +
  scale_color_manual("Algorithm:", values=c("darkslategray4", "darkseagreen3")) +
  labs(x="Variance (ms)", y="Error (ms)") +
  scale_x_continuous(breaks=c(7.5,8,8.5,9)) +
  theme_classic() +
  theme(legend.position="bottom",
    legend.key.size = unit(.2, "cm"),
    legend.title = element_text(size=10),
    legend.text = element_text(size=10),
    legend.background = element_rect(colour ="black"),
    strip.text.y = element_text(angle = 0), 
    panel.grid.minor = element_blank(),
    legend.key.width=unit(0.05,"cm"),
    strip.background = element_blank()) 
```