### 2020-03-20 voltagebudget/analysis/paper_figures_lossy.Rmd
- Split Fig 1 into Fig 1 and 2 - Misc other tweaks

---
title: "Paper figures"
output: html_notebook
---

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(readr)
library(gridExtra)
library(readr)
library(patchwork)
library(NNTbiomarker)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-mad(x)
   ymax <- m+mad(x)
   return(c(y=m, ymin=ymin, ymax=ymax))
}

mean_summary <- function(x) {
   m <- mean(x)
   return(c(y=m))
}
sample_n_groups = function(tbl, size, replace = FALSE, weight = NULL) {
  # regroup when done
  grps = tbl %>% groups %>% lapply(as.character) %>% unlist
  # check length of groups non-zero
  keep = tbl %>% summarise() %>% ungroup() %>% sample_n(size, replace, weight)
  # keep only selected groups, regroup because joins change count.
  # regrouping may be unnecessary but joins do something funky to grouping variable
  tbl %>% right_join(keep, by=grps) %>% group_by_(.dots = grps)
}

find_point <- function(x, y, point){
  i <- argmin(x, target = point)
  return(c(x[i], y[i]))
}

ms <- 1000
mV <- 1000
```

# Figure 1 - intro story
```{r, message=FALSE, fig.height=1, fig.width=3}
# Load a Poisson sampled stimulus
ref_times <- read_csv("../../data/stim1.csv")
ref_times %>% 
   mutate(ts = ts - min(ts)) -> ref_times

# Max compress the reference. Just pick a spike time
# and put it everywhere.
sync_times <- ref_times
sync_times$ts <- ref_times$ts[9] # I picked 9 for no reason.
sync_times$delta <- ref_times$ts - sync_times$ts

# Plot em....
p1 <- ref_times %>% 
  ggplot(aes(x=ts*1e3, y=ns)) +
  geom_point(alpha=0.8) +
  labs(x="Time (ms)", y="Neuron") +
  lims(x=c(0, 50)) +
  theme_classic() +
  ggtitle("Original\nmessage")

p2 <- ref_times %>% 
  ggplot(aes(x=ts*1e3, y=ns)) +
  geom_blank() +
  annotate("text", x =25, y = 20, label = "?", size=20, alpha=0.8) +
  labs(x="Time (ms)", y="Neuron") +
  lims(x=c(0, 50)) +
  theme_classic() +
  ggtitle("Optimal\ncompression")

p3 <- sync_times %>% 
  ggplot(aes(x=ts*1e3, y=ns)) +
  geom_point(alpha=0.8) +
  labs(x="Time (ms)", y="Neuron") +
  lims(x=c(0, 50)) +
  theme_classic() +
  ggtitle("Maximum\ncompression")

# Join plots into the final figure
# and save the result
# (patchwork at work)
p1 + p2 + p3 + 
  plot_annotation(tag_levels = 'a', tag_suffix = ".")
ggsave("figures/lossy_fig1.png", dpi=600)
```

# Figure 2 - error
```{r, message=FALSE, fig.height=0.85, fig.width=1}
p1 <- sync_times %>% 
  ggplot(aes(x=ts*1e3, y=ns)) +
  geom_point(alpha=0.8, size=1.2, shape=1) +
  geom_segment(aes(x=ts*ms, xend=(ts+delta)*ms, y=ns, yend=ns), 
               size=.4,
               color="darkred",
               alpha=0.8,
               arrow=arrow(length = unit(.08, "npc")),
               lineend = 'round', 
               linejoin = 'round') +
  labs(x="Time (ms)", y="Neuron") +
  lims(x=c(0, 50)) +
  theme_classic() +
  ggtitle("Compression\nerror")

p1 
ggsave("figures/lossy_fig2.png", dpi=600)
```

# Figure 3 - random sync
```{r, message=FALSE, fig.height=2.4, fig.width=2.6}
# Load a Poisson sampled stimulus
ref_times <- read_csv("../../data/stim1.csv")
ref_times %>% 
   mutate(ts = ts - min(ts)) -> ref_times

# Examples 
monte_carlo_times <- NULL
exps <- 1:9 # Only load the first 12 as examples
for(i in exps) {
  df <- read_csv(paste("../../data/exp135_mc_", as.character(i), "_target0.011_spikes.csv", sep = ""))
  df$exp <- rep(i, nrow(df))
  df$ts <- df$ts - min(df$ts)
  df$delta <- ref_times$ts - df$ts
  monte_carlo_times <- bind_rows(monte_carlo_times, df)
}

# Drop tiny deltas to clean up the visualization
monte_carlo_times$delta[monte_carlo_times$delta < 1e-3] = NA

# Errors (avg)
monte_carlo_results <- NULL
exps <- 1:100
for(i in exps) {
  df <- read_csv(paste("../../data/exp135_mc_", as.character(i), ".csv", sep = ""))
  df$exp <- rep(i, nrow(df))
  monte_carlo_results <- bind_rows(monte_carlo_results, df)
}

# Plot
p1 <- monte_carlo_times %>% 
  ggplot(aes(x=ts*ms, y=ns)) +
  geom_point(alpha=0.8, shape=1) +
  geom_segment(aes(x=ts*ms, xend=(ts+delta)*ms, y=ns, yend=ns), 
               size=.4,
               color="darkred",
               alpha=0.8,
               arrow=arrow(length = unit(.08, "npc")),
               lineend = 'round', 
               linejoin = 'round') +
  facet_wrap(~exp) +
  labs(x="Time (ms)", y="Neuron") +
  lims(x=c(0, 50)) +
  theme_classic() +
  theme(
    legend.position="none", 
    panel.border = element_rect(fill = NA, colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    strip.text.y = element_text(angle = 0)) +
  ggtitle("Random\ncompression (10%)")

p2 <- monte_carlo_results %>% 
  ggplot(aes(x=(errors - mean(errors))*ms)) +
  geom_histogram(alpha=0.4, fill="darkred") +
  # geom_histogram(alpha=0.8, bins=30, binwidth = .01) +
  geom_density(alpha=0.8, color="darkred") +
  labs(y="Count", x="Error (ms)") +
  lims(x=c(-0.25, 0.25)) +
  theme_classic()

p3 <- monte_carlo_results %>% 
  ggplot(aes(x=(obs_variances - mean(obs_variances))*ms)) +
  geom_histogram(alpha=0.4) +
  geom_density(alpha=0.8) +
  labs(y="Count", x="Synchrony (ms)") +
  lims(x=c(-0.25, 0.25)) +
  theme_classic()

p1 / (p3 + p2 + plot_spacer()) + 
  plot_layout(heights = c(3, 1)) + 
  plot_annotation(tag_levels = 'a', tag_suffix = ".")

ggsave("figures/lossy_fig3.png", dpi=600)
```