### 2020-03-20 voltagebudget/analysis/paper_figures_lossy.Rmd
Added Lossy Figures 3-6

---
title: "Paper figures"
output: html_notebook
---

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(readr)
library(gridExtra)
library(readr)
library(patchwork)
library(NNTbiomarker)

normalize <- function(x) {
    return ((x - min(x)) / (max(x) - min(x)))
}

data_summary <- function(x) {
   m <- mean(x)
   ymin <- m-mad(x)
   ymax <- m+mad(x)
   return(c(y=m, ymin=ymin, ymax=ymax))
}

mean_summary <- function(x) {
   m <- mean(x)
   return(c(y=m))
}
sample_n_groups = function(tbl, size, replace = FALSE, weight = NULL) {
  # regroup when done
  grps = tbl %>% groups %>% lapply(as.character) %>% unlist
  # check length of groups non-zero
  keep = tbl %>% summarise() %>% ungroup() %>% sample_n(size, replace, weight)
  # keep only selected groups, regroup because joins change count.
  # regrouping may be unnecessary but joins do something funky to grouping variable
  tbl %>% right_join(keep, by=grps) %>% group_by_(.dots = grps)
}

find_point <- function(x, y, point){
  i <- argmin(x, target = point)
  return(c(x[i], y[i]))
}

ms <- 1000
mV <- 1000
```

# Figure 1 and 2 - intro story, definition of error
## Figure 1 - intro
```{r, message=FALSE, fig.height=1, fig.width=3}
# Load a Poisson sampled stimulus
ref_times <- read_csv("../../data/stim1.csv")
ref_times %>% 
   mutate(ts = ts - min(ts)) -> ref_times

# Max compress the reference. Just pick a spike time
# and put it everywhere.
sync_times <- ref_times
sync_times$ts <- ref_times$ts[9] # I picked 9 for no reason.
sync_times$delta <- ref_times$ts - sync_times$ts

sync_times$delta[abs(sync_times$delta) < 1e-3] = NA

# Plot em....
p1 <- ref_times %>% 
  ggplot(aes(x=ts*1e3, y=ns)) +
  geom_point(alpha=0.8) +
  labs(x="Time (ms)", y="Neuron") +
  lims(x=c(0, 50)) +
  theme_classic() +
  ggtitle("Original\nmessage")

p2 <- ref_times %>% 
  ggplot(aes(x=ts*1e3, y=ns)) +
  geom_blank() +
  annotate("text", x =25, y = 20, label = "?", size=20, alpha=0.8) +
  labs(x="Time (ms)", y="Neuron") +
  lims(x=c(0, 50)) +
  theme_classic() +
  ggtitle("Optimal\ncompression")

p3 <- sync_times %>% 
  ggplot(aes(x=ts*1e3, y=ns)) +
  geom_point(alpha=0.8) +
  labs(x="Time (ms)", y="Neuron") +
  lims(x=c(0, 50)) +
  theme_classic() +
  ggtitle("Maximum\ncompression")

# Join plots into the final figure
# and save the result
# (patchwork at work)
p1 + p2 + p3 + 
  plot_annotation(tag_levels = 'a', tag_suffix = ".")
ggsave("figures/lossy_fig1.png", dpi=600)
```

## Figure 2 - error
```{r, message=FALSE, fig.height=0.95, fig.width=1}
p1 <- sync_times %>% 
  ggplot(aes(x=ts*1e3, y=ns)) +
  geom_point(alpha=0.8, size=1.2, shape=1) +
  geom_segment(aes(x=ts*ms, xend=(ts+delta)*ms, y=ns, yend=ns), 
               size=.4,
               color="palegreen4",
               alpha=0.99,
               # arrow=arrow(length = unit(.08, "npc")),
               lineend = 'round', 
               linejoin = 'round') +
  labs(x="Time (ms)", y="Neuron") +
  lims(x=c(0, 50)) +
  theme_classic() +
  ggtitle("Maximum\ncompression\nerror")

p1 
ggsave("figures/lossy_fig2.png", dpi=600)
```

# Figure 3 - random sync
```{r, message=FALSE, fig.height=2.4, fig.width=2.6}
# Load a Poisson sampled stimulus
ref_times <- read_csv("../../data/stim1.csv")
ref_times %>% 
   mutate(ts = ts - min(ts)) -> ref_times

# Examples 
monte_carlo_times <- NULL
exps <- 1:9 # Only load the first 12 as examples
for(i in exps) {
  df <- read_csv(paste("../../data/exp135_mc_", as.character(i), "_target0.011_spikes.csv", sep = ""))
  df$exp <- rep(i, nrow(df))
  df$ts <- df$ts - min(df$ts)
  df$delta <- ref_times$ts - df$ts
  monte_carlo_times <- bind_rows(monte_carlo_times, df)
}

# Drop tiny deltas to clean up the visualization
monte_carlo_times$delta[abs(monte_carlo_times$delta) < 2e-3] = NA

# Errors (avg)
monte_carlo_results <- NULL
exps <- 1:100
for(i in exps) {
  df <- read_csv(paste("../../data/exp135_mc_", as.character(i), ".csv", sep = ""))
  df$exp <- rep(i, nrow(df))
  monte_carlo_results <- bind_rows(monte_carlo_results, df)
}

# Plot
p1 <- monte_carlo_times %>% 
  ggplot(aes(x=ts*ms, y=ns)) +
  geom_point(alpha=0.8, shape=1) +
  geom_segment(aes(x=ts*ms, xend=(ts+delta)*ms, y=ns, yend=ns), 
               size=.6,
               color="palegreen4",
               alpha=0.99,
               # arrow=arrow(length = unit(.08, "npc")),
               lineend = 'round', 
               linejoin = 'round') +
  facet_wrap(~exp) +
  labs(x="Time (ms)", y="Neuron") +
  lims(x=c(0, 50)) +
  theme_classic() +
  theme(
    legend.position="none", 
    panel.border = element_rect(fill = NA, colour = "grey"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    strip.text.y = element_text(angle = 0)) +
  ggtitle("Random\ncompression (10%)")

p2 <- monte_carlo_results %>% 
  ggplot(aes(x=(errors - mean(errors))*ms)) +
  geom_histogram(alpha=0.4, fill="palegreen4") +
  # geom_histogram(alpha=0.8, bins=30, binwidth = .01) +
  geom_density(alpha=0.8, color="palegreen4") +
  labs(y="Count", x="Error (ms)") +
  lims(x=c(-0.25, 0.25)) +
  theme_classic()

p3 <- monte_carlo_results %>% 
  ggplot(aes(x=(obs_variances - mean(obs_variances))*ms)) +
  geom_histogram(alpha=0.4) +
  geom_density(alpha=0.8) +
  labs(y="Count", x="Synchrony (ms)") +
  lims(x=c(-0.25, 0.25)) +
  theme_classic()

p1 / (p3 + p2 + plot_spacer()) + 
  plot_layout(heights = c(3, 1)) + 
  plot_annotation(tag_levels = 'a', tag_suffix = ".")

ggsave("figures/lossy_fig3.png", dpi=600)
```

# Figure 4 - optimal algs
```{r, fig.width=1.4, fig.height=2.4, message=FALSE}
# Agls
algs <- c("max", "left", "uni", "cc")
alg_names <- c("Single cell", "left", "Population", "Information\ntheory")
exps <- c(114, 115, 116, 117)

# Load agg. results
optimal_results <- NULL
for (n in 1:4){
  df <- read_csv(paste("../../data/exp", as.character(exps[n]), "_", algs[n], ".csv", sep=""))
  df$alg <- rep(alg_names[n], nrow(df))
  optimal_results <- rbind(optimal_results, df)
}

# Load exp spikes
exps <- c(110, 111, 112, 113)
optimal_times <- NULL
for (n in 1:4){
  # df <- read_csv(paste("../../data/exp", as.character(exps[n]), "_", algs[n], "_c0.1_spikes.csv", sep=""))
  df <- read_csv(paste("../../data/exp", as.character(exps[n]), "_", algs[n], "_target0.01_spikes.csv", sep=""))
  df$ts <- df$ts - min(df$ts)
  df$delta <- ref_times$ts - df$ts
  df$alg <- rep(alg_names[n], nrow(df))
  optimal_times <- rbind(optimal_times, df)
}

# Drop lext max dev.
optimal_times %>% 
  filter(alg != "left") -> optimal_times

optimal_times$delta[abs(optimal_times$delta) < 2e-3] = NA

# Plot....
optimal_times %>% 
  ggplot(aes(x=ts*ms, y=ns)) +
  geom_point(size=1.2, alpha=0.8, shape=1) +
  geom_segment(aes(x=ts*ms, xend=(ts+delta)*ms, y=ns, yend=ns, color=alg), 
               size=.6,
               alpha=0.8,
               # arrow=arrow(length = unit(.08, "npc")),
               lineend = 'round', 
               linejoin = 'round') +
  facet_grid(alg~.) +
  scale_color_manual("Algorithm:", values=c("darkslategray4", "darkseagreen3", "chartreuse3")) +
  theme_classic() +
  labs(x="Time (ms)", y="Neuron") +
  ggtitle("Optimal\ncompression") +
  theme(
    legend.position="none", 
    panel.border = element_rect(fill = NA, colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0)) -> p1

optimal_results %>% 
  filter(alg != "left") %>%
  filter(obs_variances > 10e-3) %>% 
  ggplot(aes(x=errors*1e3, y=obs_variances*1e3, color=alg)) +
  # geom_line(size=0.1, alpha=0.99) +
  geom_line(size=1, alpha=0.99) +
  scale_color_manual("Algorithm", 
                     values=c("darkslategray4", "darkseagreen3", "chartreuse3"),
                     guide = guide_legend(nrow = 3)) +
  theme_classic() +
  coord_flip() +
  labs(x="Time (ms)", y="Neuron") +
  theme(legend.position="none",
    legend.key.size = unit(.1, "cm"),
    legend.title = element_text(size=6),
    legend.text = element_text(size=6),
    legend.background = element_rect(colour ="black")) +
  labs(x="Error (ms)", y="Synchrony (ms)") -> p2

# plot_grid <- rbind(
#   c(1, 1, 1, 1, 1),
#   c(1, 1, 1, 1, 1),
#   c(2, 2, 2, 2, NA),
#   c(2, 2, 2, 2, NA))  
# 
# grid.arrange(p1, p2, layout_matrix=plot_grid)
# rm(p1, p2, plot_grid)

p1 / p2 +
  plot_layout(heights = c(2.2, 1)) + 
  plot_annotation(tag_levels = 'a', tag_suffix = ".")
ggsave("figures/lossy_fig4.png", dpi=600)
```

# Figure 5 - osc v opt
```{r, fig.width=1.8, fig.height=3, message=FALSE}
# Load data
exp <- "118"
freqs <- c("4", "8", "12", "20", "24", "30")#, "40", "60")

exp118_results <- NULL
for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  # df %>%
    # filter(n_spikes_pop > 0) -> df

  df$freq <- rep(freq, nrow(df))
  exp118_results <- rbind(exp118_results, df)
}

exp118_results$freq <- factor(exp118_results$freq, levels=freqs)

exp <- "119"
freqs <- c("4", "8", "12", "20", "24", "30")#, "40", "60")

exp119_results <- NULL
for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  # df %>%
    # filter(n_spikes_pop > 0) -> df

  df$freq <- rep(freq, nrow(df))
  exp119_results <- rbind(exp119_results, df)
}

exp119_results$freq <- factor(exp119_results$freq, levels=freqs)

# -
exp <- "109"
freqs <- c("4", "8", "12", "20", "24", "30")#, "40", "60")

exp109_results <- NULL
for(freq in freqs){
  name <- paste("../../data/exp", exp, "_d-2e-3_f", freq, ".csv", sep="")
  df <- read_csv(paste(name, sep=""))
  # df %>%
    # filter(n_spikes_pop > 0) -> df

  df$freq <- rep(freq, nrow(df))
  exp109_results <- rbind(exp109_results, df)
}

exp119_results$freq <- factor(exp119_results$freq, levels=freqs)

# -
# alg_names <- c("Single cell", "left", "Population", "Information\ntheory")
exp118_results$alg <- rep("Single cell", nrow(exp118_results))
exp119_results$alg <- rep("Population", nrow(exp119_results))
exp109_results$alg <- rep("Information\ntheory", nrow(exp109_results))
opt_v_osc_results <- rbind(exp118_results, exp119_results, exp109_results)

# Process a bit
hospital_names <- list(
  '8'="Frequency\n\n8 Hz",
  '12'="12 Hz",
  '20'="20 Hz",
  '30'="30 Hz"
)
hospital_labeller <- function(variable,value){
  return(hospital_names[as.character(value)])
}

# Plot....
opt_v_osc_results %>% 
  filter(errors_pop > 0, freq %in% c("8")) %>%
  ggplot(aes(x=variances_pop*ms, y=errors_pop*ms)) +
  geom_point(size=0.2, alpha=0.99) + 
  geom_point(mapping=aes(x=variances_opt*ms, y=errors_opt*ms, color=alg), 
             alpha=0.99, size=0.6) +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l") +
  # facet_grid(freq~., labeller=hospital_labeller) +
  scale_color_manual("Algorithm", 
                     values=c("darkslategray4", "darkseagreen3", "chartreuse3"),
                     guide = guide_legend(nrow = 3)) +
  labs(x="Synchrony (ms)", y="Error (ms)") +
  scale_x_continuous(breaks=c(7.5,8,8.5,9)) +
  ggtitle("Optimal versus\noscillation") +
  theme_classic() +
  theme(legend.position="right",
    legend.key.size = unit(.1, "cm"),
    legend.title = element_text(size=6),
    legend.text = element_text(size=6),
    legend.background = element_rect(colour ="black"),
    strip.text.y = element_text(angle = 0), 
    panel.grid.minor = element_blank(),
    legend.key.width=unit(0.05,"cm"),
    strip.background = element_blank()) -> p1

opt_v_osc_results %>% 
  filter(errors_pop > 0, freq %in% c("8")) %>%
  ggplot(aes(x=variances_pop*ms, y=variances_opt*ms, color=alg)) +
  geom_point(size=0.8, alpha=0.3) +
  geom_abline(intercept = 0, color="black", alpha=1) +
  labs(x="Oscillation\nsynchrony (ms)", y="Optimal\nsynchrony (ms)") +
  scale_color_manual("Algorithm:", 
                     values=c("darkslategray4", "darkseagreen3", "chartreuse3"),
                     guide = "none") +
  theme_classic() +
  theme(plot.title = element_text(size = 10)) -> p2

opt_v_osc_results %>% 
  filter(errors_pop > 0, freq %in% c("8")) %>%
  ggplot(aes(x=errors_pop*ms, y=errors_opt*ms, color=alg)) +
  geom_point(size=0.8, alpha=0.5) +
  lims(x=c(0,25),y=c(0,25)) +
  geom_abline(intercept = 0, color="black", alpha=1) +
  labs(x="Oscillation\nerror (ms)", y="Optimal\nerror (ms)") +
  scale_color_manual("Algorithm:", 
                     values=c("darkslategray4", "darkseagreen3", "chartreuse3"),
                     guide = "none") +
  ggtitle(" ") +
  theme_classic() -> p3

p1 / p2 / p3 +
  plot_layout(heights = c(2, 1, 1)) + 
  plot_annotation(tag_levels = 'a', tag_suffix = ".")
ggsave("figures/lossy_fig5.png", dpi=600)
```

# Figure 6 - osc freq v optimal
```{r, fig.width=2, fig.height=3, message=FALSE}
# Plot....
opt_v_osc_results %>% 
  filter(errors_pop > 0, freq %in% c("8","12","20","30")) %>%
  ggplot(aes(x=variances_pop*ms, y=errors_pop*ms)) +
  geom_point(size=0.2, alpha=0.99) + 
  geom_point(mapping=aes(x=variances_opt*ms, y=errors_opt*ms, color=alg), 
             alpha=0.99, size=0.6) +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l") +
  facet_grid(freq~., labeller=hospital_labeller) +
  scale_color_manual("Algorithm", 
                     values=c("darkslategray4", "darkseagreen3", "chartreuse3"),
                     guide = guide_legend(nrow = 3)) +
  labs(x="Synchrony (ms)", y="Error (ms)") +
  scale_x_continuous(breaks=c(7.5,8,8.5,9)) +
  ggtitle("Optimal versus\nfrequency") +
  theme_classic() +
  theme(legend.position="right",
    legend.key.size = unit(.1, "cm"),
    legend.title = element_text(size=8),
    legend.text = element_text(size=8),
    legend.background = element_rect(colour ="black"),
    strip.text.y = element_text(angle = 0), 
    panel.grid.minor = element_blank(),
    legend.key.width=unit(0.05,"cm"),
    strip.background = element_blank()) -> p1

p1
ggsave("figures/lossy_fig6.png", dpi=600)
```

# Figure 7 - shape
- TODO?